### 2018-03-01 paper/main.tex
Add draft of paper

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[margin=1in]{geometry}

\usepackage{titlesec}

\titleformat*{\section}{\normalsize\bfseries}
\titleformat*{\subsection}{\normalsize\bfseries}
\titleformat*{\subsubsection}{\normalsize\bfseries}
\titleformat*{\paragraph}{\normalsize\bfseries}
\titleformat*{\subparagraph}{\normalsize\bfseries}

\pretolerance=10000

\twocolumn
\begin{document}

\begin{titlepage}
\title{Optimizing Pypy's Garbage Collector for Copy On Write Performance given a forky Web-Server Workload}
\author{Andrew Nelson\\CSC 550}

\maketitle

\begin{abstract}
Most research into garbage collection is focused on analyzing a single process in isolation.  This paper focuses on the garbage collection of many processes with shared memory.  We examine a deployment of a Django webserver (written in python) running in multi process mode and profile its memory usage with various garbage collection algorithms.  We then modify pypy's generational garbage collector to improve its use of copy-on-write memory.  Specifically, we make references between objects into indirect references and store those in densely packed reference tables.  That way when an object is relocated from a younger into an older generation, only the reference table needs to be updated to reflect the object's new location.  This prevents the garbage collector from unnecessarily writing to shared memory.  
\end{abstract}
\end{titlepage}

\newpage

\tableofcontents

\newpage
\pagenumbering{arabic}
\twocolumn

\section{Introduction}
Advances in interpretors for high level languages have enabled the use of high level languages in performance critical applications.  Some of the most visited websites on the Internet as of 2018 are written in high level languages such as Python, Ruby, and PHP.  A common problem with these deployments is memory usage: garbage collected languages tend to use more memory than their manually-memory-managed counterparts.  Most research into garbage collection focuses on the process in isolation.  This paper, however, examines garbage collection in a group of processes which share memory in a Copy-On-Write fashion.

\begin{enumerate}
\item We are looking at web servers.  Web servers fork a lot.
\item Forked processes share CoW memory... garbage collection is messy
\item We change garbage collector to optimize CoW by making mature pointers all indirect so mature objects don't get touched during minor collection
\end{enumerate}

\section{Background}

This section provides some background information that may help to contextualize the methods used in this paper.  Section \ref{sec:CoW} is a quick review of the Linux memory model as well as an explanation of Copy-on-Write semantics.  Sections \ref{sec:RCGC} and \ref{sec:MSGC} are a quick review of the Reference-Counting and Mark-and-Sweep algorithms used together in CPython's Garbage Collector.  Section \ref{sec:CGGC} is an overview of the Compacting Generational Garbage Collection algorithm used in PyPy and racket.  Finally, section \ref{sec:CoWGC} is an overview of how these algorithms can work against each other to increase memory used by a collection of related processes.  Those familiar with the algorithms covered in sections \ref{sec:CoW}, \ref{sec:MSGC}, and \ref{sec:CGGC} may wish to skip over those sections.  Section \ref{sec:CoWGC} is more than simple review and thus should not be skipped.

\subsection{Copy-on-Write (CoW) semantics}\label{sec:CoW}

When a process in Linux forks, a new process is spawned with an identical address space.  In order to save memory, the Linux kernel does not copy the entire address space, but rather it copies the page table of the parent process and marks all physical pages as shared and read-only \cite{VMM}.  If the physical page was already marked shared, the reference count of the page is incremented.  When either the parent or child process tries to read from a location in memory, the virtual address referenced by the process gets translated to a physical address using the page table for that process.  This happens transparently from the perspective of the process \cite{VMM}.  Since both processes can share the same physical pages in memory, the combined memory consumption of these two processes is less than the sum of the individual memory consumptions.

When either the parent or the child process attempt to write to a shared page in memory, a minor page fault occurs.  Control is shifted to the kernel which determines how to proceed.  If the physical page is still being shared, the kernel allocates a new physical page, copies the contents of the shared page into the new physical page, decrements the reference count in the shared page, and inserts the new physical page into the faulted process's page table \cite{VMM}.  If the physical page is no longer being shared (for example if the child process already wrote to the physical page in question and made their own copy) then the page is converted from a read-only shared page to a read-write private page \cite{VMM}.  In this way, memory is allocated and the process's page-table is updated in an extremely lazy manor which has proven to be very efficient in practice.

In the event that a process forks multiple times, the same mechanism still works.  A single address space can be shared among $N$ processes.  Since each page has an unsigned int dedicated to reference count, quite a few pages can refer to a single page and copying will happen as needed \cite{kernelsource,VMM}.

\subsection{Reference-Counting Garbage Collection (RCGC)}\label{sec:RCGC}

Under a Reference-Counting Garbage Collection regime, every object has an internal reference count attribute.  When new references to this object are created, the reference count attribute is incremented.  When references to this object are updated or when objects which refer to this object are collected, the reference count attribute is decremented.  When the reference count of an attribute reaches zero, it is garbage collected and all the objects to which it refers are visited accordingly \cite{GC-continuum}.

This garbage collection algorithm can be visualized as a directed graph where every node is an object in memory and where every edge is a reference from one object to another.  The reference count for each object is the the in-degree of its associated node.  When there are no edges into a node, the algorithm knows that that node (and all its out-edges) can be collected.  This scheme has advantages in that many objects can be collected as soon as they become unreachable.  However, if a group of unreachable objects form a reference cycle, their reference counts will never reach zero and so they will never be collected by a strictly reference-counted system.

This algorithm has been used extensively in CPython since python 2 and accounts for the majority of object collection in most running systems \cite{cpython3-doc}.  Whenever a reference is created, updated, or deleted, CPython visits the objects being referenced and updates their reference counts appropriately.  To account for cycles in the object graph, CPython does occasional collections using the Mark-And-Sweep algorithm.

\subsection{Mark-And-Sweep Garbage Collection (MSGC)}\label{sec:MSGC}

Mark-and-Sweep Garbage Collection can be modeled as a traversal of the object graph starting with a set of "accessible roots".  These accessible roots are objects that are being directly referenced by the program (e.g. globals, variables on the stack, intermediate values of a computation, etc.).  Starting with these nodes, the algorithm recursively visits adjacent nodes marking them as "accessible" when they are visited.  At the end of this first sweep, the algorithm has divided all objects into two sets: those which are accessible to the program (nodes tags as "accessible") and those which are not accessible to the program (nodes not tagged as "accessible").  To do the collection, the algorithm makes a pass over every object in the system.  Those not tagged as "accessible" are deleted in place \cite{GC-continuum}.

A common problem with this algorithm is that objects are deleted in-place and so the physical memory where objects reside tends to become fragmented over time.  This increases the number of physical pages in use and increases the cache miss rate.

This algorithm has been used in conjunction with Reference Counting in CPython since python 2.  While the reference counter is always running in the background during CPython execution, CPython will initiate a MSGC (usually called a "major collection event") collection once memory usage exceeds a certain threshold.  In combination, these algorithms eliminate most garbage in a timely manor while also correctly collecting cycles that form on occasion \cite{cpython3-doc}.

\subsection{Compacting Generational Garbage Collection (CGGC)}\label{sec:CGGC}

In Compacting Generational Garbage Collection, objects are divided into multiple "generations" which are each located on separate physical pages of memory.  In many cases there are 3 or 4 successively larger generations.  New objects are inserted into generation 0 (often called the "nursery").  Once generation 0 reaches a certain size, the objects in generation 0 are traversed starting with the accessible roots.  Objects in generation 0 which are still accessible are copied to generation 1.  When all the accessible objects in generation 0 have been relocated, generation 0 is cleared \cite{GC-continuum}.

When generation 1 gets larger than some threshold, it is collected in the same way and accessible objects are copied to generation 2.  Indeed when generation $N$ gets large enough, its contents are traversed, accessible objects are copied to generation $N+1$, and generation $N$ is cleared.

In most implementations there is a final generation (often called the "mature" generation) for objects that have survived all the prior collections.  There are several ways to handle this final generation - the most common being to let the garbage build up.  This implementation is not technically correct but it performs well in practice.  In PyPy, the mature generation is collected using the MSGC.  Memory fragmentation which may occur in PyPy's mature generation are possible filled when objects are copied from younger generations into mature \cite{pypy-doc}.

\begin{quotation}
"Many generational GC's are not comprehensive--they don't successfully remove all the garbage (long-lived garbage, in particular, may never get collected). But it's the fresh garbage that smells the worst. :)"

 -- C2 Wiki\cite{C2}
\end{quotation}

\subsubsection{Inter-Generational References: Always Use Protection!}

The CGGC algorithm as described above is incomplete.  It's possible for objects in one generation to refer to objects in another generation and doing a naive "accessible-roots" seeded traversal will not see these references which may cause some objects to get collected prematurely.  To prevent this we introduce a protective write barrier.

Every generation keeps a list of cross-generational references into that generation.  When an object in generation $M$ is updated to reference an object in generation $N$, that object is added to generation $N$'s cross-reference list.  When generation $N$ is collected, the garbage collector adds references in the cross-reference list to the root accessible set and those references are used when tracing which objects are in use.  When an object in generation $M$ is copied to generation $M + 1$, all the updates to that object must be updated to refer to this new location in memory \cite{GC-continuum}.

\subsection{Tragedy of the commons: Dirtying Shared Pages in CGGC}\label{sec:CoWGC}

Each of these algorithms alone serve to minimize the memory used by an individual process as it runs.  CoW allows related processes to share physical memory, and all three garbage collection algorithms reduce the number of physical pages needed by each process.  However, when used in conjunction we often see that the memory performance of a collection of related processes suffers.

As a thought experiment, consider a forking webserver written in python running atop a standard CPython interpretor.  When a request comes in, the process forks to spawn a worker thread which begins processing the request.  The parent process has much in memory the child process can reference during its computation.

As the child process continues, it will surely create or update a reference to some existing constant data structure.  When this happens, the reference count field in the referenced object must be updated.  This triggers a copy-on-write because even though none of of the objects stored on the page changed, the meta-data that the garbage collector stores on the page has changed.  In this way, reference counting breaks shared pages which can increase memory used by the system.

When a CPython major collection event occurs, the MSGC algorithm will start a full heap traversal that marks objects as "accessible" or "not-accessible".  Because the garbage collector stores this information adjacent to the objects in memory, the shared pages containing these objects get written to which triggers a copy-on-write and the page sharing breaks.  In this way, a single major garbage collection event can greatly increase memory used by the system.

This problem is somewhat alleviated when using a generational garbage collector such as is included in PyPy because we can copy objects out of a shared page without writing to it, however when we do the copy, those inter-generational references must be updated.  Doing a minor collection of generation 1 may force writes to generations 2, 3, 4, etc, which breaks the sharing of those pages.

While these algorithms are successful at reducing the memory usage of an individual process, when examining a collection or related processes they do poorly at reducing total combined memory usage.  The next section is dedicated to modifying CGGC to solve this problem.

\section{Design}

The default garbage collector in PyPy is called "minimark" and it is a classic Generational Garbage Collector with 2 generations \cite{pypy-doc}.  New objects are created in the nursery and are moved to gen1 on minor collection events.

\begin{enumerate}
	\item Currently when mature objects refer to old objects, that reference is stored in a special "write barrier" so their references are counted
    \item When young object moves from young to mature generation, the objects which refer to it get updated in-place
    \item We are going to put all references between objects in a reference table... every lookup will be a double lookup but every update will be a single edit to the table
\end{enumerate}

\section{Implementation}
\begin{enumerate}
	\item Give a summary of which structures and functions actually changed
    \item Where did I store this information?  How is it accessed in code?  
    \item maybe some pseudocode snippets
\end{enumerate}

\section{Methodologies}
\begin{enumerate}
  \item Ngnx + uWSGI + Django + pypy in multi-process mode
  \item Pick some website that uses a lot of memory... thebluealliance.com is open source and operates on django and caches a lot of info locally
  \item Generate a bunch of fake traffic 
  \item Monitor running code
  \begin{enumerate}
  	\item Total memory of all processes
    \item Average memory of each process
    \item shared memory per process
    \item num page faults
    \item cache miss rate?
    \item pages served per second (did it go up or down?)
    \item How many times did major and minor garbage collection happen?
    \item How many times did this inter-generation optimization actually occur?
  \end{enumerate}
\end{enumerate}

\section{Analysis}
\begin{enumerate}
  \item Did we actually save any memory?  How much on average?  Were there any patterns?   
  \item How often did old objects actually refer to young objects?  
  \item How often did minor collection happen?  Major collection?
  \item How much CPU overhead was there in following that extra pointer every time?  Can I even measure this?
  \item How many pages did we serve per second?  Did it go up or down?
\end{enumerate}

\section{Future Work}
\begin{enumerate}
	\item What other changes could I make to improve CoW performance?  Lemme think.
\end{enumerate}

\section{Related work}
\begin{enumerate}
  \item Ruby's Copy-On-Write optimized mark-and-sweep garbage collector
  \item Instagram's CPython patch for python's mark-and-sweep garbage collector
\end{enumerate}

\newpage
\onecolumn
\nocite{*}
\bibliographystyle{IEEEannot}
\bibliography{bibliography.bib}

\end{document}

### 2018-03-01 paper/main.tex
Wrote related work section

@@
 
-\usepackage{titlesec}
+%\usepackage{titlesec}
 
-\titleformat*{\section}{\normalsize\bfseries}
-\titleformat*{\subsection}{\normalsize\bfseries}
-\titleformat*{\subsubsection}{\normalsize\bfseries}
-\titleformat*{\paragraph}{\normalsize\bfseries}
-\titleformat*{\subparagraph}{\normalsize\bfseries}
+%\titleformat*{\section}{\normalsize\bfseries}
+%\titleformat*{\subsection}{\normalsize\bfseries}
+%\titleformat*{\subsubsection}{\normalsize\bfseries}
+%\titleformat*{\paragraph}{\normalsize\bfseries}
+%\titleformat*{\subparagraph}{\normalsize\bfseries}
 
@@
 	\item Ngnx + uWSGI + Django + pypy in multi-process mode
-  \item Pick some website that uses a lot of memory... thebluealliance.com is open source and operates on django and caches a lot of info locally
-  \item Generate a bunch of fake traffic 
+	\item Use thebluealliance.com as a test load becaues it caches a lot of data and does some non-trivial computation
+	\item Generate a bunch of fake traffic using httperf
 	\item Monitor running code
 	  \begin{enumerate}
+        \item How long did it take to handle 10,000 requests
+        \item What was the average time-to-last-byte?
 		\item Total memory of all processes
@@
   \section{Related work}
-\begin{enumerate}
-  \item Ruby's Copy-On-Write optimized mark-and-sweep garbage collector
-  \item Instagram's CPython patch for python's mark-and-sweep garbage collector
-\end{enumerate}
+
+  This paper is very much inspired by recent updates in CPython's and Ruby MRI's garbage collectors which were updated in 2017 and 2014 respectively.
+
+  \subsection{Copy-On-Write optimization in Ruby MRI's MSGC}
+  Ruby 2.0 contains an update to the GC that improved CoW performance.
+Ruby's GC is a classic CGGC with a young generation and an old generation.
+In the old algorithm, every object contained a "visited" bit that would note whether the object had visited by the tracer.
+In the new GC algorithm, these "visited" bits are stored outside the objects being traced.
+In this way, the objects themselves don't need to be updated during a GC event.
+
+  The optimization described in this paper is different in that we focus on preserving old objects when objects are copied from eden into the older generation.
+
+  \subsection{Copy-On-Write Friendly Garbage Collection in CPython at Instagram}
+  A 2017 article from Instagram's Engineering series describes changes to CPython's garbage collector made at instagram aimed at improving memory performance in their web-tier servers.
+The experimentation section of this paper is heavily inspired by the test setup used at Instagram.
+
+  The optimization implemented at instagram was to have the programmer manually mark certain objects as "GC-Frozen".
+Objects that are GC-Frozen are not touched by the garbage collector and so the data in and around them is not touched.
+Because Instagram's specific use case involves a lot of static objects that are shared between processes, this worked effectively.
 

### 2018-03-01 paper/main.tex
Added intro.  Fixed formatting

@@
 A common problem with these deployments is memory usage: garbage collected languages tend to use more memory than their manually-memory-managed counterparts.
+In order for these popular websites to continue to scale up, the tools on which they are developed need to get more efficient.
+Webservers need to use less memory.
+
 Most research into garbage collection focuses on the process in isolation.
 This paper, however, examines garbage collection in a group of processes which share memory in a Copy-On-Write fashion.
+In a webserver environment where worker processes are often forked repeatedly during operation, this situation is very common.
+Workers forked from the master process initially share some memory with the master and so the memory usage of the system as a whole is directly related to how much memory can be shared between processes.
+This is the type of environment that many of today's largest websites like Instagram\cite{dismissing_garbage} operate with and so we believe that improvements made in this area can be used to improve performance in the field as a whole.
 
-  \begin{enumerate}
-	\item We are looking at web servers.  Web servers fork a lot.
-	\item Forked processes share CoW memory... garbage collection is messy
-	\item We change garbage collector to optimize CoW by making mature pointers all indirect so mature objects don't get touched during minor collection
-  \end{enumerate}
+To accomplish this goal, this paper examines an optimization in PyPy's Generational Garbage Collector which is designed to reduce the number of shared pages that get dirtied during the copying phase of a minor collection.
+Section \ref{sec:background} explains various topics relating to the optimization and builds up to a demonstration intended to create some intuition into how this optimziation is designed to work.
+Section \ref{sec:design} explains the nitty gritty details of the optimization and builds intuition as to which data structures need to be changed.
+Section \ref{sec:implementation} summarizes the source changes made to PyPy's garbage collector in order to implement this.
+Section \ref{sec:method} talks about how these changes were tested and section \ref{sec:analysis} examines the results of these tests and explains them.
+Finally, section \ref{sec:relatedwork} talks about other optimizations that have been made to Garbage Collectors to optimize for Copy-On-write behavior and explains how they differ from this research.
 
-  \section{Background}
+\section{Background}\label{sec:background}
 
@@
 
-  \section{Design}
+\section{Design}\label{sec:design}
 
@@
 
-  \section{Implementation}
+\section{Implementation}\label{sec:implementation}
 \begin{enumerate}
@@
 
-  \section{Methodologies}
+\section{Methodologies}\label{sec:method}
 \begin{enumerate}
     \item Ngnx + uWSGI + Django + pypy in multi-process mode
-	\item Use thebluealliance.com as a test load becaues it caches a lot of data and does some non-trivial computation
+    \item Use thebluealliance.com as a test load because it caches a lot of data and does some non-trivial computation
     \item Generate a bunch of fake traffic using httperf
@@
 
-  \section{Analysis}
+\section{Analysis}\label{sec:analysis}
 \begin{enumerate}
@@
 
-  \section{Future Work}
+\section{Future Work}\label{sec:futurework}
 \begin{enumerate}
@@
 
-  \section{Related work}
+\section{Related work}\label{sec:relatedwork}
 
@@
 
-  The optimization described in this paper is different in that we focus on preserving old objects when objects are copied from eden into the older generation.
+The optimization described in this paper is different in that we focus on preserving old objects when objects are copied from Eden into the older generation.
 
 \subsection{Copy-On-Write Friendly Garbage Collection in CPython at Instagram}
-  A 2017 article from Instagram's Engineering series describes changes to CPython's garbage collector made at instagram aimed at improving memory performance in their web-tier servers.
+A 2017 article from Instagram's Engineering series describes changes to CPython's garbage collector made at Instagram aimed at improving memory performance in their web-tier servers.
 The experimentation section of this paper is heavily inspired by the test setup used at Instagram.
 
-  The optimization implemented at instagram was to have the programmer manually mark certain objects as "GC-Frozen".
+The optimization implemented at Instagram was to have the programmer manually mark certain objects as "GC-Frozen".
 Objects that are GC-Frozen are not touched by the garbage collector and so the data in and around them is not touched.

### 2018-03-21 test_env/mysite/crunch/apps.py
Running home-build pypy with a django environment that makes a lot of garbage

@@ -1,0 +1,5 @@
+from django.apps import AppConfig
+
+
+class CrunchConfig(AppConfig):
+    name = 'crunch'

### 2018-03-21 test_env/mysite/crunch/views.py

@@ -1,0 +1,11 @@
+from django.http import HttpResponse
+
+def index(request):
+    lists = []
+    strs = []
+    for i in range(16000):
+        lists.append([])
+        for j in range(40):
+            strs.append(' ' * 8)
+
+    return HttpResponse("Hello, world.")

### 2018-03-21 test_env/mysite/mysite/settings.py

@@ -119,3 +119,10 @@ USE_TZ = True
 
 STATIC_URL = '/static/'
 STATIC_ROOT = os.path.join(BASE_DIR, "static/")
+
+lists = []
+strs = []
+for i in range(16000):
+    lists.append(['poop'] * i)
+    for j in range(40):
+        strs.append('lol' * 8)

### 2018-03-21 test_env/mysite/mysite/urls.py

@@ -13,9 +13,10 @@ Including another URLconf
     1. Import the include() function: from django.conf.urls import url, include
     2. Add a URL to urlpatterns:  url(r'^blog/', include('blog.urls'))
 """
-from django.conf.urls import url
+from django.conf.urls import url, include
 from django.contrib import admin
 
 urlpatterns = [
     url(r'^admin/', admin.site.urls),
+    url(r'^', include('crunch.urls')),
 ]

### 2018-03-21 test_env/pypy-src/pypy/interpreter/astcompiler/test/test_astbuilder.py

@@ -1,0 +1,1252 @@
+# -*- coding: utf-8 -*-
+import random
+import string
+import sys
+import py
+from pypy.interpreter.baseobjspace import W_Root
+from pypy.interpreter.pyparser import pyparse
+from pypy.interpreter.pyparser.error import SyntaxError
+from pypy.interpreter.astcompiler.astbuilder import ast_from_node
+from pypy.interpreter.astcompiler import ast, consts
+
+
+class TestAstBuilder:
+
+    def setup_class(cls):
+        cls.parser = pyparse.PythonParser(cls.space)
+
+    def get_ast(self, source, p_mode="exec"):
+        info = pyparse.CompileInfo("<test>", p_mode,
+                                   consts.CO_FUTURE_WITH_STATEMENT)
+        tree = self.parser.parse_source(source, info)
+        ast_node = ast_from_node(self.space, tree, info)
+        return ast_node
+
+    def get_first_expr(self, source):
+        mod = self.get_ast(source)
+        assert len(mod.body) == 1
+        expr = mod.body[0]
+        assert isinstance(expr, ast.Expr)
+        return expr.value
+
+    def get_first_stmt(self, source):
+        mod = self.get_ast(source)
+        assert len(mod.body) == 1
+        return mod.body[0]
+
+    def test_top_level(self):
+        mod = self.get_ast("hi = 32")
+        assert isinstance(mod, ast.Module)
+        body = mod.body
+        assert len(body) == 1
+
+        mod = self.get_ast("hi", p_mode="eval")
+        assert isinstance(mod, ast.Expression)
+        assert isinstance(mod.body, ast.expr)
+
+        mod = self.get_ast("x = 23", p_mode="single")
+        assert isinstance(mod, ast.Interactive)
+        assert len(mod.body) == 1
+        mod = self.get_ast("x = 23; y = 23; b = 23", p_mode="single")
+        assert isinstance(mod, ast.Interactive)
+        assert len(mod.body) == 3
+        for stmt in mod.body:
+            assert isinstance(stmt, ast.Assign)
+        assert mod.body[-1].targets[0].id == "b"
+
+        mod = self.get_ast("x = 23; y = 23; b = 23")
+        assert isinstance(mod, ast.Module)
+        assert len(mod.body) == 3
+        for stmt in mod.body:
+            assert isinstance(stmt, ast.Assign)
+
+    def test_print(self):
+        pri = self.get_first_stmt("print x")
+        assert isinstance(pri, ast.Print)
+        assert pri.dest is None
[…]

### 2018-03-21 test_env/pypy-src/pypy/interpreter/test/test_unicodehelper.py

@@ -1,0 +1,49 @@
+import pytest
+import struct
+from pypy.interpreter.unicodehelper import (
+    encode_utf8, decode_utf8, unicode_encode_utf_32_be)
+
+class FakeSpace:
+    pass
+
+def test_encode_utf8():
+    space = FakeSpace()
+    assert encode_utf8(space, u"abc") == "abc"
+    assert encode_utf8(space, u"\u1234") == "\xe1\x88\xb4"
+    assert encode_utf8(space, u"\ud800") == "\xed\xa0\x80"
+    assert encode_utf8(space, u"\udc00") == "\xed\xb0\x80"
+    # for the following test, go to lengths to avoid CPython's optimizer
+    # and .pyc file storage, which collapse the two surrogates into one
+    c = u"\udc00"
+    assert encode_utf8(space, u"\ud800" + c) == "\xf0\x90\x80\x80"
+
+def test_decode_utf8():
+    space = FakeSpace()
+    assert decode_utf8(space, "abc") == u"abc"
+    assert decode_utf8(space, "\xe1\x88\xb4") == u"\u1234"
+    assert decode_utf8(space, "\xed\xa0\x80") == u"\ud800"
+    assert decode_utf8(space, "\xed\xb0\x80") == u"\udc00"
+    got = decode_utf8(space, "\xed\xa0\x80\xed\xb0\x80")
+    assert map(ord, got) == [0xd800, 0xdc00]
+    got = decode_utf8(space, "\xf0\x90\x80\x80")
+    assert map(ord, got) == [0x10000]
+
+@pytest.mark.parametrize('unich', [u"\ud800", u"\udc80"])
+def test_utf32_surrogates(unich):
+    assert (unicode_encode_utf_32_be(unich, 1, None) ==
+            struct.pack('>i', ord(unich)))
+    with pytest.raises(UnicodeEncodeError):
+        unicode_encode_utf_32_be(unich, 1, None, allow_surrogates=False)
+
+    def replace_with(ru, rs):
+        def errorhandler(errors, enc, msg, u, startingpos, endingpos):
+            if errors == 'strict':
+                raise UnicodeEncodeError(enc, u, startingpos, endingpos, msg)
+            return ru, rs, endingpos
+        return unicode_encode_utf_32_be(
+            u"<%s>" % unich, 3, None,
+            errorhandler, allow_surrogates=False)
+
[…]

### 2018-03-21 test_env/pypy-src/pypy/module/_sre/interp_sre.py

@@ -1,0 +1,658 @@
+import sys
+from pypy.interpreter.baseobjspace import W_Root
+from pypy.interpreter.typedef import GetSetProperty, TypeDef
+from pypy.interpreter.typedef import interp_attrproperty, interp_attrproperty_w
+from pypy.interpreter.typedef import make_weakref_descr
+from pypy.interpreter.gateway import interp2app, unwrap_spec, WrappedDefault
+from pypy.interpreter.error import OperationError, oefmt
+from rpython.rlib.rarithmetic import intmask
+from rpython.rlib import jit
+from rpython.rlib.rstring import StringBuilder, UnicodeBuilder
+
+# ____________________________________________________________
+#
+# Constants and exposed functions
+
+from rpython.rlib.rsre import rsre_core
+from rpython.rlib.rsre.rsre_char import CODESIZE, MAXREPEAT, getlower, set_unicode_db
+
+
+@unwrap_spec(char_ord=int, flags=int)
+def w_getlower(space, char_ord, flags):
+    return space.newint(getlower(char_ord, flags))
+
+
+def w_getcodesize(space):
+    return space.newint(CODESIZE)
+
+# use the same version of unicodedb as the standard objspace
+import pypy.objspace.std.unicodeobject
+set_unicode_db(pypy.objspace.std.unicodeobject.unicodedb)
+
+# ____________________________________________________________
+#
+
+
+def slice_w(space, ctx, start, end, w_default):
+    if 0 <= start <= end:
+        if isinstance(ctx, rsre_core.BufMatchContext):
+            return space.newbytes(ctx._buffer.getslice(start, end, 1,
+                                                        end-start))
+        if isinstance(ctx, rsre_core.StrMatchContext):
+            return space.newbytes(ctx._string[start:end])
+        elif isinstance(ctx, rsre_core.UnicodeMatchContext):
+            return space.newunicode(ctx._unicodestr[start:end])
+        else:
+            # unreachable
+            raise SystemError
+    return w_default
+
+
+@jit.look_inside_iff(lambda ctx, num_groups: jit.isconstant(num_groups))
+def do_flatten_marks(ctx, num_groups):
+    # Returns a list of RPython-level integers.
+    # Unlike the app-level groups() method, groups are numbered from 0
+    # and the returned list does not start with the whole match range.
+    if num_groups == 0:
+        return None
+    result = [-1] * (2 * num_groups)
+    mark = ctx.match_marks
+    while mark is not None:
[…]

### 2018-03-21 test_env/pypy-src/pypy/module/_sre/test/test_app_sre.py

@@ -1,0 +1,1021 @@
+"""Regular expression tests specific to _sre.py and accumulated during TDD."""
+
+import os
+import py
+from py.test import raises, skip
+from pypy.interpreter.gateway import app2interp_temp
+
+
+def init_app_test(cls, space):
+    cls.w_s = space.appexec(
+        [space.wrap(os.path.realpath(os.path.dirname(__file__)))],
+        """(this_dir):
+        import sys
+        # Uh-oh, ugly hack
+        sys.path.insert(0, this_dir)
+        try:
+            import support_test_app_sre
+            return support_test_app_sre
+        finally:
+            sys.path.pop(0)
+        """)
+
+
+class AppTestSrePy:
+    def test_magic(self):
+        import _sre, sre_constants
+        assert sre_constants.MAGIC == _sre.MAGIC
+
+    def test_codesize(self):
+        import _sre
+        assert _sre.getcodesize() == _sre.CODESIZE
+
+
+class AppTestSrePattern:
+    def setup_class(cls):
+        # This imports support_test_sre as the global "s"
+        init_app_test(cls, cls.space)
+
+    def test_copy(self):
+        # copy support is disabled by default in _sre.c
+        import re
+        p = re.compile("b")
+        raises(TypeError, p.__copy__)        # p.__copy__() should raise
+        raises(TypeError, p.__deepcopy__)    # p.__deepcopy__() should raise
+
+    def test_creation_attributes(self):
+        import re
+        pattern_string = "(b)l(?P<g>a)"
+        p = re.compile(pattern_string, re.I | re.M)
+        assert pattern_string == p.pattern
+        assert re.I | re.M == p.flags
+        assert 2 == p.groups
+        assert {"g": 2} == p.groupindex
+
+    def test_repeat_minmax_overflow(self):
+        import re
+        string = "x" * 100000
+        assert re.match(r".{%d}" % (self.s.MAXREPEAT - 1), string) is None
+        assert re.match(r".{,%d}" % (self.s.MAXREPEAT - 1), string).span() == (0, 100000)
+        assert re.match(r".{%d,}?" % (self.s.MAXREPEAT - 1), string) is None
+        raises(OverflowError, re.compile, r".{%d}" % self.s.MAXREPEAT)
+        raises(OverflowError, re.compile, r".{,%d}" % self.s.MAXREPEAT)
+        raises(OverflowError, re.compile, r".{%d,}?" % self.s.MAXREPEAT)
+
+    def test_match_none(self):
+        import re
[…]

### 2018-03-21 test_env/pypy-src/pypy/module/cpyext/__init__.py

@@ -1,0 +1,81 @@
+from pypy.interpreter.mixedmodule import MixedModule
+from pypy.module.cpyext.state import State
+from pypy.module.cpyext import api
+
+class Module(MixedModule):
+    interpleveldefs = {
+        'load_module': 'api.load_extension_module',
+        'is_cpyext_function': 'interp_cpyext.is_cpyext_function',
+    }
+
+    appleveldefs = {
+    }
+
+    atexit_funcs = []
+
+    def startup(self, space):
+        space.fromcache(State).startup(space)
+        method = pypy.module.cpyext.typeobject.get_new_method_def(space)
+        # the w_self argument here is a dummy, the only thing done with w_obj
+        # is call space.type on it
+        w_obj = pypy.module.cpyext.methodobject.W_PyCFunctionObject(space, method, space.w_None)
+        space.appexec([space.type(w_obj)], """(methodtype):
+            from pickle import Pickler
+            Pickler.dispatch[methodtype] = Pickler.save_global
+        """)
+
+    def register_atexit(self, function):
+        if len(self.atexit_funcs) >= 32:
+            raise ValueError("cannot register more than 32 atexit functions")
+        self.atexit_funcs.append(function)
+
+    def shutdown(self, space):
+        for func in self.atexit_funcs:
+            func()
+
+
+# import these modules to register api functions by side-effect
+import pypy.module.cpyext.pyobject
+import pypy.module.cpyext.boolobject
+import pypy.module.cpyext.floatobject
+import pypy.module.cpyext.modsupport
+import pypy.module.cpyext.pythonrun
+import pypy.module.cpyext.pyerrors
+import pypy.module.cpyext.typeobject
+import pypy.module.cpyext.object
+import pypy.module.cpyext.bytesobject
+import pypy.module.cpyext.bytearrayobject
+import pypy.module.cpyext.tupleobject
+import pypy.module.cpyext.setobject
+import pypy.module.cpyext.dictobject
+import pypy.module.cpyext.intobject
+import pypy.module.cpyext.longobject
+import pypy.module.cpyext.listobject
+import pypy.module.cpyext.sequence
+import pypy.module.cpyext.buffer
+import pypy.module.cpyext.bufferobject
+import pypy.module.cpyext.eval
+import pypy.module.cpyext.import_
+import pypy.module.cpyext.mapping
+import pypy.module.cpyext.iterator
[…]

### 2018-03-21 test_env/pypy-src/pypy/module/cpyext/longobject.py

@@ -1,0 +1,240 @@
+from rpython.rtyper.lltypesystem import lltype, rffi
+from pypy.module.cpyext.api import (
+    cpython_api, PyObject, build_type_checkers_flags, Py_ssize_t,
+    CONST_STRING, ADDR, CANNOT_FAIL)
+from pypy.objspace.std.longobject import W_LongObject
+from pypy.interpreter.error import OperationError
+from pypy.module.cpyext.intobject import PyInt_AsUnsignedLongMask
+from rpython.rlib.rbigint import rbigint
+
+PyLong_Check, PyLong_CheckExact = build_type_checkers_flags("Long")
+
+@cpython_api([lltype.Signed], PyObject)
+def PyLong_FromLong(space, val):
+    """Return a new PyLongObject object from v, or NULL on failure."""
+    return space.newlong(val)
+
+@cpython_api([Py_ssize_t], PyObject)
+def PyLong_FromSsize_t(space, val):
+    """Return a new PyLongObject object from a C Py_ssize_t, or
+    NULL on failure.
+    """
+    return space.newlong(val)
+
+@cpython_api([rffi.SIZE_T], PyObject)
+def PyLong_FromSize_t(space, val):
+    """Return a new PyLongObject object from a C size_t, or NULL on
+    failure.
+    """
+    return space.newlong_from_rarith_int(val)
+
+@cpython_api([rffi.LONGLONG], PyObject)
+def PyLong_FromLongLong(space, val):
+    """Return a new PyLongObject object from a C long long, or NULL
+    on failure."""
+    return space.newlong_from_rarith_int(val)
+
+@cpython_api([rffi.ULONG], PyObject)
+def PyLong_FromUnsignedLong(space, val):
+    """Return a new PyLongObject object from a C unsigned long, or
+    NULL on failure."""
+    return space.newlong_from_rarith_int(val)
+
+@cpython_api([rffi.ULONGLONG], PyObject)
+def PyLong_FromUnsignedLongLong(space, val):
+    """Return a new PyLongObject object from a C unsigned long long,
+    or NULL on failure."""
+    return space.newlong_from_rarith_int(val)
+
+@cpython_api([PyObject], rffi.ULONG, error=-1)
+def PyLong_AsUnsignedLong(space, w_long):
+    """
+    Return a C unsigned long representation of the contents of pylong.
+    If pylong is greater than ULONG_MAX, an OverflowError is
+    raised."""
+    try:
[…]

### 2018-03-21 test_env/pypy-src/pypy/module/cpyext/pyerrors.py

@@ -1,0 +1,428 @@
+import os
+
+from rpython.rtyper.lltypesystem import rffi, lltype
+from pypy.interpreter.error import OperationError, oefmt
+from pypy.interpreter import pytraceback
+from pypy.module.cpyext.api import cpython_api, CANNOT_FAIL, CONST_STRING
+from pypy.module.exceptions.interp_exceptions import W_RuntimeWarning
+from pypy.module.cpyext.pyobject import (
+    PyObject, PyObjectP, make_ref, from_ref, Py_DecRef)
+from pypy.module.cpyext.state import State
+from pypy.module.cpyext.import_ import PyImport_Import
+from rpython.rlib import rposix, jit
+
+@cpython_api([PyObject, PyObject], lltype.Void)
+def PyErr_SetObject(space, w_type, w_value):
+    """This function is similar to PyErr_SetString() but lets you specify an
+    arbitrary Python object for the "value" of the exception."""
+    state = space.fromcache(State)
+    state.set_exception(OperationError(w_type, w_value))
+
+@cpython_api([PyObject, CONST_STRING], lltype.Void)
+def PyErr_SetString(space, w_type, message_ptr):
+    message = rffi.charp2str(message_ptr)
+    PyErr_SetObject(space, w_type, space.newtext(message))
+
+@cpython_api([PyObject], lltype.Void, error=CANNOT_FAIL)
+def PyErr_SetNone(space, w_type):
+    """This is a shorthand for PyErr_SetObject(type, Py_None)."""
+    PyErr_SetObject(space, w_type, space.w_None)
+
+@cpython_api([], PyObject, result_borrowed=True)
+def PyErr_Occurred(space):
+    state = space.fromcache(State)
+    if state.operror is None:
+        return None
+    return state.operror.w_type     # borrowed ref
+
+@cpython_api([], lltype.Void)
+def PyErr_Clear(space):
+    state = space.fromcache(State)
+    state.clear_exception()
+
+@cpython_api([PyObject], PyObject)
+def PyExceptionInstance_Class(space, w_obj):
+    return space.type(w_obj)
+
+@cpython_api([PyObjectP, PyObjectP, PyObjectP], lltype.Void)
+def PyErr_Fetch(space, ptype, pvalue, ptraceback):
+    """Retrieve the error indicator into three variables whose addresses are passed.
+    If the error indicator is not set, set all three variables to NULL.  If it is
+    set, it will be cleared and you own a reference to each object retrieved.  The
[…]

### 2018-03-21 test_env/pypy-src/pypy/module/cpyext/test/test_cpyext.py

@@ -1,0 +1,928 @@
+import sys
+
+import pytest
+
+from pypy.tool.cpyext.extbuild import SystemCompilationInfo, HERE
+from pypy.interpreter.gateway import unwrap_spec, interp2app
+from pypy.interpreter.error import OperationError
+from rpython.rtyper.lltypesystem import lltype
+from pypy.module.cpyext import api
+from pypy.module.cpyext.api import cts
+from pypy.module.cpyext.pyobject import from_ref
+from pypy.module.cpyext.state import State
+from rpython.tool import leakfinder
+from rpython.rlib import rawrefcount
+from rpython.tool.udir import udir
+
+only_pypy ="config.option.runappdirect and '__pypy__' not in sys.builtin_module_names"
+
+@api.cpython_api([], api.PyObject)
+def PyPy_Crash1(space):
+    1/0
+
+@api.cpython_api([], lltype.Signed, error=-1)
+def PyPy_Crash2(space):
+    1/0
+
+@api.cpython_api([api.PyObject], api.PyObject, result_is_ll=True)
+def PyPy_Noop(space, pyobj):
+    return pyobj
+
+class TestApi:
+    def test_signature(self):
+        common_functions = api.FUNCTIONS_BY_HEADER[api.pypy_decl]
+        assert 'PyModule_Check' in common_functions
+        assert common_functions['PyModule_Check'].argtypes == [api.PyObject]
+
+
+class SpaceCompiler(SystemCompilationInfo):
+    """Extension compiler for regular (untranslated PyPy) mode"""
+    def __init__(self, space, *args, **kwargs):
+        self.space = space
+        SystemCompilationInfo.__init__(self, *args, **kwargs)
+
+    def load_module(self, mod, name):
+        space = self.space
+        api.load_extension_module(space, mod, name)
+        return space.getitem(
+            space.sys.get('modules'), space.wrap(name))
+
+
+def get_cpyext_info(space):
+    from pypy.module.imp.importing import get_so_extension
+    state = space.fromcache(State)
+    api_library = state.api_lib
+    if sys.platform == 'win32':
+        libraries = [api_library]
+        # '%s' undefined; assuming extern returning int
+        compile_extra = ["/we4013"]
+        # prevent linking with PythonXX.lib
+        w_maj, w_min = space.fixedview(space.sys.get('version_info'), 5)[:2]
+        link_extra = ["/NODEFAULTLIB:Python%d%d.lib" %
[…]

### 2018-03-21 test_env/pypy-src/pypy/module/cpyext/test/test_eval.py

@@ -1,0 +1,422 @@
+import sys
+import os
+import pytest
+from rpython.rtyper.lltypesystem import rffi, lltype
+from pypy.module.cpyext.test.test_cpyext import AppTestCpythonExtensionBase
+from pypy.module.cpyext.test.test_api import BaseApiTest, raises_w
+from pypy.module.cpyext.object import PyObject_Size, PyObject_GetItem
+from pypy.module.cpyext.pythonrun import Py_AtExit
+from pypy.module.cpyext.eval import (
+    Py_single_input, Py_file_input, Py_eval_input, PyCompilerFlags,
+    PyEval_CallObjectWithKeywords, PyObject_CallObject, PyEval_EvalCode,
+    PyRun_SimpleString, PyRun_String, PyRun_StringFlags, PyRun_File,
+    PyEval_GetBuiltins, PyEval_GetLocals, PyEval_GetGlobals,
+    _PyEval_SliceIndex)
+from pypy.module.cpyext.api import (
+    c_fopen, c_fclose, c_fileno, Py_ssize_tP, is_valid_fd)
+from pypy.interpreter.gateway import interp2app
+from pypy.interpreter.error import OperationError
+from pypy.interpreter.astcompiler import consts
+from rpython.tool.udir import udir
+
+class TestEval(BaseApiTest):
+    def test_eval(self, space):
+        w_l, w_f = space.fixedview(space.appexec([], """():
+        l = []
+        def f(arg1, arg2):
+            l.append(arg1)
+            l.append(arg2)
+            return len(l)
+        return l, f
+        """))
+
+        w_t = space.newtuple([space.wrap(1), space.wrap(2)])
+        w_res = PyEval_CallObjectWithKeywords(space, w_f, w_t, None)
+        assert space.int_w(w_res) == 2
+        assert space.len_w(w_l) == 2
+        w_f = space.appexec([], """():
+            def f(*args, **kwds):
+                assert isinstance(kwds, dict)
+                assert 'xyz' in kwds
+                return len(kwds) + len(args) * 10
+            return f
+            """)
+        w_t = space.newtuple([space.w_None, space.w_None])
+        w_d = space.newdict()
+        space.setitem(w_d, space.wrap("xyz"), space.wrap(3))
+        w_res = PyEval_CallObjectWithKeywords(space, w_f, w_t, w_d)
+        assert space.int_w(w_res) == 21
+
+    def test_call_object(self, space):
[…]

### 2018-03-21 test_env/pypy-src/pypy/module/cpyext/test/test_methodobject.py

@@ -1,0 +1,136 @@
+from pypy.module.cpyext.test.test_api import BaseApiTest
+from pypy.module.cpyext.test.test_cpyext import AppTestCpythonExtensionBase
+from pypy.module.cpyext.methodobject import PyMethodDef
+from pypy.module.cpyext.api import ApiFunction
+from pypy.module.cpyext.pyobject import PyObject, make_ref, Py_DecRef
+from pypy.module.cpyext.methodobject import (
+    PyDescr_NewMethod, PyCFunction)
+from rpython.rtyper.lltypesystem import rffi, lltype
+
+class AppTestMethodObject(AppTestCpythonExtensionBase):
+    def test_call_METH(self):
+        mod = self.import_extension('MyModule', [
+            ('getarg_O', 'METH_O',
+             '''
+             Py_INCREF(args);
+             return args;
+             '''
+             ),
+            ('getarg_NO', 'METH_NOARGS',
+             '''
+             if(args) {
+                 Py_INCREF(args);
+                 return args;
+             }
+             else {
+                 Py_INCREF(Py_None);
+                 return Py_None;
+             }
+             '''
+             ),
+            ('getarg_OLD', 'METH_OLDARGS',
+             '''
+             if(args) {
+                 Py_INCREF(args);
+                 return args;
+             }
+             else {
+                 Py_INCREF(Py_None);
+                 return Py_None;
+             }
+             '''
+             ),
+            ('isCFunction', 'METH_O',
+             '''
+             if(PyCFunction_Check(args)) {
+                 PyCFunctionObject* func = (PyCFunctionObject*)args;
+                 return PyString_FromString(func->m_ml->ml_name);
+             }
+             else {
+                 Py_RETURN_FALSE;
+             }
+             '''
+             ),
+            ('getModule', 'METH_O',
+             '''
+             if(PyCFunction_Check(args)) {
+                 PyCFunctionObject* func = (PyCFunctionObject*)args;
+                 Py_INCREF(func->m_module);
+                 return func->m_module;
+             }
+             else {
+                 Py_RETURN_FALSE;
+             }
+             '''
+             ),
+            ('isSameFunction', 'METH_O',
+             '''
+             PyCFunction ptr = PyCFunction_GetFunction(args);
+             if (!ptr) return NULL;
+             if (ptr == (PyCFunction)MyModule_getarg_O)
+                 Py_RETURN_TRUE;
[…]

### 2018-03-21 test_env/pypy-src/pypy/module/cpyext/test/test_object.py

@@ -1,0 +1,571 @@
+import pytest
+
+from pypy.module.cpyext.test.test_api import BaseApiTest, raises_w
+from pypy.module.cpyext.test.test_cpyext import AppTestCpythonExtensionBase
+from rpython.rtyper.lltypesystem import rffi, lltype
+from pypy.module.cpyext.api import (
+    Py_LT, Py_LE, Py_NE, Py_EQ, Py_GE, Py_GT)
+from pypy.module.cpyext.object import (
+    PyObject_IsTrue, PyObject_Not, PyObject_GetAttrString,
+    PyObject_DelAttrString, PyObject_GetAttr, PyObject_DelAttr,
+    PyObject_GetItem,
+    PyObject_IsInstance, PyObject_IsSubclass, PyObject_AsFileDescriptor,
+    PyObject_Hash, PyObject_Cmp, PyObject_Unicode
+)
+
+class TestObject(BaseApiTest):
+    def test_IsTrue(self, space, api):
+        assert api.PyObject_IsTrue(space.wrap(1.0)) == 1
+        assert api.PyObject_IsTrue(space.wrap(False)) == 0
+        assert api.PyObject_IsTrue(space.wrap(0)) == 0
+
+    def test_Not(self, space, api):
+        assert api.PyObject_Not(space.wrap(False)) == 1
+        assert api.PyObject_Not(space.wrap(0)) == 1
+        assert api.PyObject_Not(space.wrap(True)) == 0
+        assert api.PyObject_Not(space.wrap(3.14)) == 0
+
+    def test_exception(self, space, api):
+        w_obj = space.appexec([], """():
+            class C:
+                def __nonzero__(self):
+                    raise ValueError
+            return C()""")
+
+        with raises_w(space, ValueError):
+            PyObject_IsTrue(space, w_obj)
+        with raises_w(space, ValueError):
+            PyObject_Not(space, w_obj)
+
+    def test_HasAttr(self, space, api):
+        hasattr_ = lambda w_obj, name: api.PyObject_HasAttr(w_obj,
+                                                            space.wrap(name))
+        assert hasattr_(space.wrap(''), '__len__')
+        assert hasattr_(space.w_int, '__eq__')
+        assert not hasattr_(space.w_int, 'nonexistingattr')
+
+        buf = rffi.str2charp('__len__')
+        assert api.PyObject_HasAttrString(space.w_bytes, buf)
[…]

### 2018-03-21 test_env/pypy-src/pypy/module/cpyext/test/test_tupleobject.py

@@ -1,0 +1,211 @@
+import py
+
+from pypy.module.cpyext.pyobject import (
+    PyObject, PyObjectP, make_ref, from_ref, incref)
+from pypy.module.cpyext.test.test_api import BaseApiTest, raises_w
+from pypy.module.cpyext.test.test_cpyext import AppTestCpythonExtensionBase
+from rpython.rtyper.lltypesystem import rffi, lltype
+from rpython.rlib.debug import FatalError
+from pypy.module.cpyext.tupleobject import (
+    PyTupleObject, PyTuple_Check, PyTuple_SetItem, PyTuple_Size)
+
+
+class TestTupleObject(BaseApiTest):
+
+    def test_tupleobject(self, space):
+        assert not PyTuple_Check(space, space.w_None)
+        with raises_w(space, SystemError):
+            incref(space, space.w_None)
+            PyTuple_SetItem(space, space.w_None, 0, space.w_None)
+        atuple = space.newtuple([space.wrap(0), space.wrap(1),
+                                 space.wrap('yay')])
+        assert PyTuple_Size(space, atuple) == 3
+        with raises_w(space, SystemError):
+            PyTuple_Size(space, space.newlist([]))
+
+    def test_tuple_realize_refuses_nulls(self, space, api):
+        py_tuple = api.PyTuple_New(1)
+        py.test.raises(FatalError, from_ref, space, py_tuple)
+        api.Py_DecRef(py_tuple)
+
+    def test_tuple_resize(self, space, api):
+        w_42 = space.wrap(42)
+        w_43 = space.wrap(43)
+        w_44 = space.wrap(44)
+        ar = lltype.malloc(PyObjectP.TO, 1, flavor='raw')
+
+        py_tuple = api.PyTuple_New(3)
+        # inside py_tuple is an array of "PyObject *" items which each hold
+        # a reference
+        rffi.cast(PyTupleObject, py_tuple).c_ob_item[0] = make_ref(space, w_42)
+        rffi.cast(PyTupleObject, py_tuple).c_ob_item[1] = make_ref(space, w_43)
+        ar[0] = py_tuple
+        api._PyTuple_Resize(ar, 2)
+        w_tuple = from_ref(space, ar[0])
+        assert space.int_w(space.len(w_tuple)) == 2
+        assert space.int_w(space.getitem(w_tuple, space.wrap(0))) == 42
[…]

### 2018-03-21 test_env/pypy-src/pypy/module/posix/test/test_posix2.py

@@ -1,0 +1,1298 @@
+
+# -*- coding: utf-8 -*-
+
+from __future__ import with_statement
+from pypy.objspace.std import StdObjSpace
+from rpython.tool.udir import udir
+from pypy.tool.pytest.objspace import gettestobjspace
+from pypy import pypydir
+from rpython.translator.c.test.test_extfunc import need_sparse_files
+from rpython.rlib import rposix
+import os
+import py
+import sys
+import signal
+
+USEMODULES = ['binascii', 'posix', 'struct', 'time']
+if os.name != 'nt':
+    USEMODULES += ['fcntl']
+else:
+    # On windows, os.popen uses the subprocess module
+    USEMODULES += ['_rawffi', 'thread', 'signal']
+
+def setup_module(mod):
+    mod.space = gettestobjspace(usemodules=USEMODULES)
+    mod.path = udir.join('posixtestfile.txt')
+    mod.path.write("this is a test")
+    mod.path2 = udir.join('test_posix2-')
+    pdir = udir.ensure('posixtestdir', dir=True)
+    pdir.join('file1').write("test1")
+    os.chmod(str(pdir.join('file1')), 0600)
+    pdir.join('file2').write("test2")
+    pdir.join('another_longer_file_name').write("test3")
+    mod.pdir = pdir
+    if sys.platform == 'darwin':
+        # see issue https://bugs.python.org/issue31380
+        unicode_dir = udir.ensure('fixc5x9fier.txt', dir=True)
+        file_name = 'cafxe9'
+    else:
+        unicode_dir = udir.ensure('fi\xc5\x9fier.txt', dir=True)
+        file_name = 'caf\xe9'
+    unicode_dir.join('somefile').write('who cares?')
+    unicode_dir.join(file_name).write('who knows?')
+    mod.unicode_dir = unicode_dir
+
+    # in applevel tests, os.stat uses the CPython os.stat.
+    # Be sure to return times with full precision
+    # even when running on top of CPython 2.4.
+    os.stat_float_times(True)
+
+    # Initialize sys.filesystemencoding
+    # space.call_method(space.getbuiltinmodule('sys'), 'getfilesystemencoding')
+
+
+
+GET_POSIX = "(): import %s as m ; return m" % os.name
+
+
+class AppTestPosix:
+    spaceconfig = {'usemodules': USEMODULES}
+
+    def setup_class(cls):
+        space = cls.space
[…]

### 2018-03-21 test_env/pypy-src/pypy/module/sys/version.py

@@ -1,0 +1,112 @@
+"""
+Version numbers exposed by PyPy through the 'sys' module.
+"""
+import os
+from rpython.rlib import compilerinfo
+from pypy.interpreter import gateway
+
+#XXX # the release serial 42 is not in range(16)
+CPYTHON_VERSION            = (2, 7, 13, "final", 42)
+#XXX # sync CPYTHON_VERSION with patchlevel.h, package.py
+CPYTHON_API_VERSION        = 1013   #XXX # sync with include/modsupport.h
+
+PYPY_VERSION               = (5, 10, 0, "final", 0)    #XXX # sync patchlevel.h
+
+
+import pypy
+pypydir = pypy.pypydir
+pypyroot = os.path.dirname(pypydir)
+del pypy
+from rpython.tool.version import get_repo_version_info
+
+import time as t
+gmtime = t.gmtime()
+date = t.strftime("%b %d %Y", gmtime)
+time = t.strftime("%H:%M:%S", gmtime)
+del t
+
+# ____________________________________________________________
+
+app = gateway.applevel('''
+"NOT_RPYTHON"
+from _structseq import structseqtype, structseqfield
+class version_info:
+    __metaclass__ = structseqtype
+
+    major        = structseqfield(0, "Major release number")
+    minor        = structseqfield(1, "Minor release number")
+    micro        = structseqfield(2, "Patch release number")
+    releaselevel = structseqfield(3,
+                       "'alpha', 'beta', 'candidate', or 'release'")
+    serial       = structseqfield(4, "Serial release number")
+''')
+
+def get_api_version(space):
+    return space.newint(CPYTHON_API_VERSION)
+
+def get_version_info(space):
+    w_version_info = app.wget(space, "version_info")
+    # run at translation time
+    return space.call_function(w_version_info, space.wrap(CPYTHON_VERSION))
+
+def _make_version_template(PYPY_VERSION=PYPY_VERSION):
+    ver = "%d.%d.%d" % (PYPY_VERSION[0], PYPY_VERSION[1], PYPY_VERSION[2])
+    if PYPY_VERSION[3] != "final":
+        ver = ver + "-%s%d" %(PYPY_VERSION[3], PYPY_VERSION[4])
+    template = "%d.%d.%d (%s, %s, %s)\n[PyPy %s with %%s]" % (
+        CPYTHON_VERSION[0],
+        CPYTHON_VERSION[1],
+        CPYTHON_VERSION[2],
[…]

### 2018-03-21 test_env/pypy-src/pypy/objspace/std/listobject.py

@@ -1,0 +1,2171 @@
+"""The builtin list implementation
+
+Lists optimize their storage by holding certain primitive datatypes in
+unwrapped form. For more information:
+
+http://morepypy.blogspot.com/2011/10/more-compact-lists-with-list-strategies.html
+
+"""
+
+import operator
+import sys
+
+from rpython.rlib import debug, jit, rerased
+from rpython.rlib.listsort import make_timsort_class
+from rpython.rlib.objectmodel import (
+    import_from_mixin, instantiate, newlist_hint, resizelist_hint, specialize)
+from rpython.rlib import longlong2float
+from rpython.tool.sourcetools import func_with_new_name
+
+from pypy.interpreter.baseobjspace import W_Root
+from pypy.interpreter.error import OperationError, oefmt
+from pypy.interpreter.gateway import (
+    WrappedDefault, applevel, interp2app, unwrap_spec)
+from pypy.interpreter.signature import Signature
+from pypy.interpreter.typedef import TypeDef
+from pypy.interpreter.miscutils import StringSort
+from pypy.objspace.std.bytesobject import W_BytesObject
+from pypy.objspace.std.floatobject import W_FloatObject
+from pypy.objspace.std.intobject import W_IntObject
+from pypy.objspace.std.iterobject import (
+    W_FastListIterObject, W_ReverseSeqIterObject)
+from pypy.objspace.std.sliceobject import (
+    W_SliceObject, normalize_simple_slice, unwrap_start_stop)
+from pypy.objspace.std.tupleobject import W_AbstractTupleObject
+from pypy.objspace.std.unicodeobject import W_UnicodeObject
+from pypy.objspace.std.util import get_positive_index, negate
+
+__all__ = ['W_ListObject', 'make_range_list', 'make_empty_list_with_size']
+
+
+UNROLL_CUTOFF = 5
+
+
+def make_range_list(space, start, step, length):
+    if length <= 0:
+        strategy = space.fromcache(EmptyListStrategy)
+        storage = strategy.erase(None)
+    elif start == 0 and step == 1:
+        strategy = space.fromcache(SimpleRangeListStrategy)
+        storage = strategy.erase((length,))
+    else:
+        strategy = space.fromcache(RangeListStrategy)
+        storage = strategy.erase((start, step, length))
+    return W_ListObject.from_storage_and_strategy(space, storage, strategy)
+
+
+def make_empty_list(space):
[…]

### 2018-03-21 test_env/pypy-src/pypy/tool/release/force-builds.py

@@ -1,0 +1,88 @@
+#!/usr/bin/python
+
+"""
+Force the PyPy buildmaster to run a builds on all builders that produce
+nightly builds for a particular branch.
+
+Taken from http://twistedmatrix.com/trac/browser/sandbox/exarkun/force-builds.py
+
+modified by PyPy team
+"""
+
+import os, sys, urllib, subprocess
+
+from twisted.internet import reactor, defer
+from twisted.python import log
+from twisted.web import client
+from twisted.web.error import PageRedirect
+
+BUILDERS = [
+    'own-linux-x86-32',
+    'own-linux-x86-64',
+#    'own-linux-armhf',
+    'own-win-x86-32',
+    'own-linux-s390x',
+#    'own-macosx-x86-32',
+    'pypy-c-jit-linux-x86-32',
+    'pypy-c-jit-linux-x86-64',
+#    'pypy-c-jit-freebsd-9-x86-64',
+    'pypy-c-jit-macosx-x86-64',
+    'pypy-c-jit-win-x86-32',
+    'pypy-c-jit-linux-s390x',
+    'build-pypy-c-jit-linux-armhf-raspbian',
+    'build-pypy-c-jit-linux-armel',
+]
+
+def get_user():
+    if sys.platform == 'win32':
+        return os.environ['USERNAME']
+    else:
+        import pwd
+        return pwd.getpwuid(os.getuid())[0]
+
+def main(branch, server, user):
+    #XXX: handle release tags
+    #XXX: handle validity checks
+    lock = defer.DeferredLock()
+    requests = []
+    def ebList(err):
+        if err.check(PageRedirect) is not None:
+            return None
+        log.err(err, "Build force failure")
+
+    for builder in BUILDERS:
+        print 'Forcing', builder, '...'
+        url = "http://" + server + "/builders/" + builder + "/force"
+        args = [
+            ('username', user),
+            ('revision', ''),
+            ('forcescheduler', 'Force Scheduler'),
+            ('submit', 'Force Build'),
+            ('branch', branch),
+            ('comments', "Forced by command line script")]
+        url = url + '?' + '&'.join([k + '=' + urllib.quote(v) for (k, v) in args])
+        requests.append(
+            lock.run(client.getPage, url, followRedirect=False).addErrback(ebList))
+
+    d = defer.gatherResults(requests)
+    d.addErrback(log.err)
+    d.addCallback(lambda ign: reactor.stop())
+    reactor.run()
[…]

### 2018-03-21 test_env/pypy-src/rpython/rlib/objectmodel.py

@@ -1,0 +1,1031 @@
+"""
+This file defines utilities for manipulating objects in an
+RPython-compliant way.
+"""
+
+from __future__ import absolute_import
+
+import sys
+import types
+import math
+import inspect
+from collections import OrderedDict
+
+from rpython.tool.sourcetools import rpython_wrapper, func_with_new_name
+from rpython.rtyper.extregistry import ExtRegistryEntry
+from rpython.flowspace.specialcase import register_flow_sc
+from rpython.flowspace.model import Constant
+
+# specialize is a decorator factory for attaching _annspecialcase_
+# attributes to functions: for example
+#
+# f._annspecialcase_ = 'specialize:memo' can be expressed with:
+# @specialize.memo()
+# def f(...
+#
+# f._annspecialcase_ = 'specialize:arg(0)' can be expressed with:
+# @specialize.arg(0)
+# def f(...
+#
+
+
+class _Specialize(object):
+    def memo(self):
+        """ Specialize the function based on argument values.  All arguments
+        have to be either constants or PBCs (i.e. instances of classes with a
+        _freeze_ method returning True).  The function call is replaced by
+        just its result, or in case several PBCs are used, by some fast
+        look-up of the result.
+        """
+        def decorated_func(func):
+            func._annspecialcase_ = 'specialize:memo'
+            return func
+        return decorated_func
+
+    def arg(self, *args):
+        """ Specialize the function based on the values of given positions
+        of arguments.  They must be compile-time constants in order to work.
+
+        There will be a copy of provided function for each combination
+        of given arguments on positions in args (that can lead to
+        exponential behavior!).
+        """
+        def decorated_func(func):
+            func._annspecialcase_ = 'specialize:arg' + self._wrap(args)
+            return func
+
+        return decorated_func
+
+    def arg_or_var(self, *args):
+        """ Same as arg, but additionally allow for a 'variable' annotation,
+        that would simply be a situation where designated arg is not
+        a constant
+        """
+        def decorated_func(func):
+            func._annspecialcase_ = 'specialize:arg_or_var' + self._wrap(args)
+            return func
+
+        return decorated_func
+
+    def argtype(self, *args):
+        """ Specialize function based on types of arguments on given positions.
+
+        There will be a copy of provided function for each combination
+        of given arguments on positions in args (that can lead to
+        exponential behavior!).
+        """
+        def decorated_func(func):
[…]

### 2018-03-21 test_env/pypy-src/rpython/rlib/rposix.py

@@ -1,0 +1,2576 @@
+import os
+import sys
+import errno
+from rpython.annotator.model import s_Str0
+from rpython.rtyper.lltypesystem.rffi import CConstant, CExternVariable, INT
+from rpython.rtyper.lltypesystem import lltype, ll2ctypes, rffi
+from rpython.rtyper.tool import rffi_platform
+from rpython.rlib import debug, jit, rstring, rthread, types
+from rpython.rlib._os_support import (
+    _CYGWIN, _MACRO_ON_POSIX, UNDERSCORE_ON_WIN32, _WIN32,
+    _prefer_unicode, _preferred_traits)
+from rpython.rlib.objectmodel import (
+    specialize, enforceargs, register_replacement_for, NOT_CONSTANT)
+from rpython.rlib.rarithmetic import intmask, widen
+from rpython.rlib.signature import signature
+from rpython.tool.sourcetools import func_renamer
+from rpython.translator.platform import platform
+from rpython.translator.tool.cbuild import ExternalCompilationInfo
+
+
+if _WIN32:
+    from rpython.rlib import rwin32
+    from rpython.rlib.rwin32file import make_win32_traits
+
+
+class CConstantErrno(CConstant):
+    # these accessors are used when calling get_errno() or set_errno()
+    # on top of CPython
+    def __getitem__(self, index):
+        assert index == 0
+        try:
+            return ll2ctypes.TLS.errno
+        except AttributeError:
+            raise ValueError("no C function call occurred so far, "
+                             "errno is undefined")
+    def __setitem__(self, index, value):
+        assert index == 0
+        ll2ctypes.TLS.errno = value
+
+if os.name == 'nt':
+    if platform.name == 'msvc':
+        includes=['errno.h','stdio.h']
+    else:
+        includes=['errno.h','stdio.h', 'stdint.h']
+    separate_module_sources =['''
+        /* Lifted completely from CPython 3.3 Modules/posix_module.c */
+        #include <malloc.h> /* for _msize */
+        typedef struct {
+            intptr_t osfhnd;
+            char osfile;
+        } my_ioinfo;
+        extern __declspec(dllimport) char * __pioinfo[];
+        #define IOINFO_L2E 5
+        #define IOINFO_ARRAY_ELTS   (1 << IOINFO_L2E)
+        #define IOINFO_ARRAYS 64
[…]

### 2018-03-21 test_env/pypy-src/rpython/rlib/rvmprof/cintf.py

@@ -1,0 +1,236 @@
+import platform as host_platform
+import py
+import sys
+import shutil
+from rpython.tool.udir import udir
+from rpython.tool.version import rpythonroot
+from rpython.rtyper.lltypesystem import lltype, llmemory, rffi
+from rpython.translator.tool.cbuild import ExternalCompilationInfo
+from rpython.rtyper.tool import rffi_platform as platform
+from rpython.rlib import rthread, jit
+from rpython.rlib.objectmodel import we_are_translated
+from rpython.config.translationoption import get_translation_config
+from rpython.jit.backend import detect_cpu
+
+class VMProfPlatformUnsupported(Exception):
+    pass
+
+# vmprof works only on x86 for now
+IS_SUPPORTED = detect_cpu.autodetect().startswith('x86')
+if sys.platform == 'win32':
+    IS_SUPPORTED = False
+
+ROOT = py.path.local(rpythonroot).join('rpython', 'rlib', 'rvmprof')
+SRC = ROOT.join('src')
+SHARED = SRC.join('shared')
+BACKTRACE = SHARED.join('libbacktrace')
+
+def make_eci():
+    if make_eci.called:
+        raise ValueError("make_eci() should be called at most once")
+    #
+    compile_extra = ['-DRPYTHON_VMPROF']
+    separate_module_files = [
+        SHARED.join('symboltable.c'),
+        SHARED.join('vmprof_unix.c')
+    ]
+    if sys.platform.startswith('linux'):
+        separate_module_files += [
+           BACKTRACE.join('atomic.c'),
+           BACKTRACE.join('backtrace.c'),
+           BACKTRACE.join('state.c'),
+           BACKTRACE.join('elf.c'),
+           BACKTRACE.join('dwarf.c'),
+           BACKTRACE.join('fileline.c'),
+           BACKTRACE.join('mmap.c'),
+           BACKTRACE.join('mmapio.c'),
+           BACKTRACE.join('posix.c'),
+           BACKTRACE.join('sort.c'),
+        ]
+        _libs = ['dl']
+        compile_extra += ['-DVMPROF_UNIX']
+        compile_extra += ['-DVMPROF_LINUX']
+    elif sys.platform == 'win32':
+        compile_extra += ['-DVMPROF_WINDOWS']
+        separate_module_files = [SHARED.join('vmprof_win.c')]
+        _libs = []
+    else:
+        # Guessing a BSD-like Unix platform
[…]

### 2018-03-21 test_env/pypy-src/rpython/rlib/rwin32.py

@@ -1,0 +1,573 @@
+""" External functions accessing the win32 api.
+Common types, functions from core win32 libraries, such as kernel32
+"""
+
+import os
+import errno
+
+from rpython.rlib.rposix_environ import make_env_impls
+from rpython.rtyper.tool import rffi_platform
+from rpython.tool.udir import udir
+from rpython.translator.tool.cbuild import ExternalCompilationInfo
+from rpython.translator.platform import CompilationError
+from rpython.rtyper.lltypesystem import lltype, rffi
+from rpython.rlib.rarithmetic import intmask
+from rpython.rlib import jit
+
+# This module can be imported on any platform,
+# but most symbols are not usable...
+WIN32 = os.name == "nt"
+
+if WIN32:
+    eci = ExternalCompilationInfo(
+        includes = ['windows.h', 'stdio.h', 'stdlib.h'],
+        libraries = ['kernel32'],
+        )
+else:
+    eci = ExternalCompilationInfo()
+
+class CConfig:
+    _compilation_info_ = eci
+
+    if WIN32:
+        DWORD_PTR = rffi_platform.SimpleType("DWORD_PTR", rffi.LONG)
+        WORD = rffi_platform.SimpleType("WORD", rffi.UINT)
+        DWORD = rffi_platform.SimpleType("DWORD", rffi.UINT)
+        BOOL = rffi_platform.SimpleType("BOOL", rffi.LONG)
+        BYTE = rffi_platform.SimpleType("BYTE", rffi.UCHAR)
+        WCHAR = rffi_platform.SimpleType("WCHAR", rffi.UCHAR)
+        INT = rffi_platform.SimpleType("INT", rffi.INT)
+        LONG = rffi_platform.SimpleType("LONG", rffi.LONG)
+        PLONG = rffi_platform.SimpleType("PLONG", rffi.LONGP)
+        LPVOID = rffi_platform.SimpleType("LPVOID", rffi.INTP)
+        LPCVOID = rffi_platform.SimpleType("LPCVOID", rffi.VOIDP)
+        LPSTR = rffi_platform.SimpleType("LPSTR", rffi.CCHARP)
+        LPCSTR = rffi_platform.SimpleType("LPCSTR", rffi.CCHARP)
+        LPWSTR = rffi_platform.SimpleType("LPWSTR", rffi.CWCHARP)
[…]

### 2018-03-21 test_env/pypy-src/rpython/rlib/test/test_rarithmetic.py

@@ -1,0 +1,688 @@
+from rpython.rtyper.test.tool import BaseRtypingTest
+from rpython.rtyper.test.test_llinterp import interpret
+from rpython.rlib.rarithmetic import *
+from rpython.rlib.rstring import ParseStringError, ParseStringOverflowError
+from hypothesis import given, strategies, assume
+import sys
+import py
+
+maxint_mask = (sys.maxint*2 + 1)
+machbits = 0
+i = 1
+l = 1L
+while i == l and type(i) is int:
+    i *= 2
+    l *= 2
+    machbits += 1
+#print machbits
+
+
+class Test_r_int:
+    def test__add__(self):
+        self.binary_test(lambda x, y: x + y, includes_floats=True)
+    def test__sub__(self):
+        self.binary_test(lambda x, y: x - y, includes_floats=True)
+    def test__mul__(self):
+        self.binary_test(lambda x, y: x * y, includes_floats=True)
+        x = 3; y = [2]
+        assert x*y == r_int(x)*y
+        assert y*x == y*r_int(x)
+    def test__div__(self):
+        self.binary_test(lambda x, y: x // y)
+    def test__mod__(self):
+        self.binary_test(lambda x, y: x % y)
+    def test__divmod__(self):
+        self.binary_test(divmod)
+    def test__lshift__(self):
+        self.binary_test(lambda x, y: x << y, (1, 2, 3))
+    def test__rshift__(self):
+        self.binary_test(lambda x, y: x >> y, (1, 2, 3))
+    def test__or__(self):
+        self.binary_test(lambda x, y: x | y)
+    def test__and__(self):
+        self.binary_test(lambda x, y: x & y)
+    def test__xor__(self):
+        self.binary_test(lambda x, y: x ^ y)
+    def test__neg__(self):
+        self.unary_test(lambda x: -x)
+    def test__pos__(self):
+        self.unary_test(lambda x: +x)
+    def test__invert__(self):
+        self.unary_test(lambda x: ~x)
+    def test__pow__(self):
+        self.binary_test(lambda x, y: x**y, (2, 3))
+        self.binary_test(lambda x, y: pow(x, y, 42L), (2, 3, 5, 1000))
+
+    def unary_test(self, f):
+        for arg in (-10, -1, 0, 3, 12345):
+            res = f(arg)
+            cmp = f(r_int(arg))
+            assert res == cmp
+
+    def binary_test(self, f, rargs=None, includes_floats=False):
+        if not rargs:
[…]

### 2018-03-21 test_env/pypy-src/rpython/rlib/test/test_runicode.py

@@ -1,0 +1,962 @@
+# -*- coding: utf-8 -*-
+
+import py
+import sys, random
+import struct
+from rpython.rlib import runicode
+
+from hypothesis import given, settings, strategies
+
+
+def test_unichr():
+    assert runicode.UNICHR(0xffff) == u'\uffff'
+    if runicode.MAXUNICODE > 0xffff:
+        if sys.maxunicode < 0x10000:
+            assert runicode.UNICHR(0x10000) == u'\ud800\udc00'
+        else:
+            assert runicode.UNICHR(0x10000) == u'\U00010000'
+    else:
+        py.test.raises(ValueError, runicode.UNICHR, 0x10000)
+    py.test.raises(TypeError, runicode.UNICHR, 'abc')
+
+
+def test_ord():
+    assert runicode.ORD('a') == 97
+    assert runicode.ORD(u'a') == 97
+    assert runicode.ORD(u'\uffff') == 0xffff
+    if runicode.MAXUNICODE > 0xffff:
+        if sys.maxunicode < 0x10000:
+            assert runicode.ORD(u'\ud800\udc00') == 0x10000
+        else:
+            assert runicode.ORD(u'\U00010000') == 0x10000
+    else:
+        py.test.raises(TypeError, runicode.ORD, u'\ud800\udc00')
+    py.test.raises(TypeError, runicode.ORD, 'abc')
+
+
+class UnicodeTests(object):
+    def typeequals(self, x, y):
+        assert x == y
+        assert type(x) is type(y)
+
+    def getdecoder(self, encoding, look_for_py3k=False):
+        prefix = "py3k_" if look_for_py3k else ""
+        return getattr(runicode, "%sstr_decode_%s" %
+                                 (prefix, encoding.replace("-", "_")))
+
+    def getencoder(self, encoding):
+        return getattr(runicode,
+                       "unicode_encode_%s" % encoding.replace("-", "_"))
+
+    def checkdecode(self, s, encoding):
+        decoder = self.getdecoder(encoding)
+        try:
+            if isinstance(s, str):
+                trueresult = s.decode(encoding)
+            else:
+                trueresult = s
+                s = s.encode(encoding)
+        except LookupError as e:
+            py.test.skip(e)
+        result, consumed = decoder(s, len(s), 'strict', final=True)
+        assert consumed == len(s)
+        self.typeequals(trueresult, result)
+
+    def checkencode(self, s, encoding):
+        encoder = self.getencoder(encoding)
+        try:
[…]

### 2018-03-21 test_env/run_test.py

@@ -1,0 +1,102 @@
+from pprint import pprint
+import subprocess
+from collections import defaultdict
+from httperfpy import Httperf, HttperfParser
+import matplotlib.pyplot as plt
+import json
+from threading import Thread
+
+measurements = defaultdict(lambda: {})
+num_requests = 0
+
+def get_pids():
+    res = subprocess.check_output(
+            "ps -u root -ae | grep uwsgi | grep -v grep",
+            shell=True)
+    processes = res.decode().strip().split('\n')
+    processes = [x.strip() for x in processes]
+    processes = [x.split(' ')[0] for x in processes]
+    processes = [int(x) for x in processes]
+    return processes
+
+def get_stat(pid, stat):
+    res = subprocess.check_output(
+            "cat /proc/{}/smaps | grep {}".format(pid, stat),
+            shell=True
+    ).decode().strip().split('\n')
+    res = [int(x.strip().split(' ')[-2]) for x in res]
+    return sum(res)
+
+def measure_pid(pid):
+    shared_memory = get_stat(pid, "Shared_Clean") + get_stat(pid, "Shared_Dirty")
+    private_memory = get_stat(pid, "Private_Clean") + get_stat(pid, "Private_Dirty")
+
+    global num_requests
+    measurements[pid][num_requests] = {
+            'shared': shared_memory,
+            'private': private_memory,
+            'total': shared_memory + private_memory,
+    }
+
+def measure_all_pids():
+    for pid in get_pids():
+        try:
+            measure_pid(pid)
+        except:
+            print "Couldn't get measurement for {} at {}".format(pid, num_requests)
+
+def test_load(requests=30):
+    global num_requests
+    num_requests += requests * 2
+
+    def attach_load():
+        perf = Httperf(server="localhost", port=4000,
+                       num_conns=requests)
+        httperf_result_string = perf.run()
+
+        results = HttperfParser.parse(httperf_result_string)
+
+    a = Thread(target=attach_load)
+    b = Thread(target=attach_load)
+    a.start()
+    b.start()
+    a.join()
+    b.join()
+
+def plot_results():
+    fig, (shared, private) = plt.subplots(2, 1)
+
+    Xs = []
+    shared_Ys = []
+    private_Ys = []
+    for worker in measurements.values():
+        for x, vals in worker.items():
+            Xs.append(x)
+            shared_Ys.append(vals['shared'])
[…]

### 2018-03-22 test_env/mysite/crunch/views.py
Working set

@@ -1,11 +1,24 @@
 from django.http import HttpResponse
 
+class LargeObject():
+    def __init__(self):
+        lists = []
+        strs = []
+        for i in range(1600):
+            lists.append([])
+            for j in range(40):
+                strs.append(' ' * 8)
+        self.lists = lists
+        self.strs = strs
+
+glob_a = None
+glob_b = None
+glob_c = None
+
 def index(request):
-    lists = []
-    strs = []
-    for i in range(16000):
-        lists.append([])
-        for j in range(40):
-            strs.append(' ' * 8)
+    global glob_a
+    global glob_b
+    global glob_c
+    glob_a, glob_b, glob_c = LargeObject(), glob_a, glob_b
 
     return HttpResponse("Hello, world.")

### 2018-03-22 test_env/pypy-withpatch/pypy/module/_codecs/test/test_codecs.py

@@ -1,0 +1,784 @@
+import sys
+
+class AppTestCodecs:
+    spaceconfig = {
+        "usemodules": ['unicodedata', 'struct', 'binascii'],
+    }
+
+    def test_register_noncallable(self):
+        import _codecs
+        raises(TypeError, _codecs.register, 1)
+
+    def test_bigU_codecs(self):
+        u = u'\U00010001\U00020002\U00030003\U00040004\U00050005'
+        for encoding in ('utf-8', 'utf-16', 'utf-16-le', 'utf-16-be',
+                         'utf-32', 'utf-32-le', 'utf-32-be',
+                         'raw_unicode_escape',
+                         'unicode_escape', 'unicode_internal'):
+            assert unicode(u.encode(encoding),encoding) == u
+
+    def test_ucs4(self):
+        x = u'\U00100000'
+        y = x.encode("raw-unicode-escape").decode("raw-unicode-escape")
+        assert x == y
+
+    def test_named_unicode(self):
+        assert unicode('\\N{SPACE}','unicode-escape') == u" "
+        raises( UnicodeDecodeError, unicode,'\\N{SPACE','unicode-escape')
+        raises( UnicodeDecodeError, unicode,'\\NSPACE}','unicode-escape')
+        raises( UnicodeDecodeError, unicode,'\\NSPACE','unicode-escape')
+        raises( UnicodeDecodeError, unicode,'\\N','unicode-escape')
+        assert  unicode('\\N{SPACE}\\N{SPACE}','unicode-escape') == u"  "
+        assert  unicode('\\N{SPACE}a\\N{SPACE}','unicode-escape') == u" a "
+        assert "\\N{foo}xx".decode("unicode-escape", "ignore") == u"xx"
+        assert 1 <= len(u"\N{CJK UNIFIED IDEOGRAPH-20000}") <= 2
+
+    def test_literals(self):
+        raises(SyntaxError, eval, 'u\'\\Uffffffff\'')
+
+    def test_insecure_pickle(self):
+        import pickle
+        insecure = ["abc", "2 + 2", # not quoted
+                    #"'abc' + 'def'", # not a single quoted string
+                    "'abc", # quote is not closed
+                    "'abc\"", # open quote and close quote don't match
+                    "'abc'   ?", # junk after close quote
+                    "'\\'", # trailing backslash
+                    # some tests of the quoting rules
+                    #"'abc\"\''",
[…]

### 2018-03-22 test_env/pypy-withpatch/pypy/module/cpyext/include/patchlevel.h

@@ -1,0 +1,51 @@
+
+/* Newfangled version identification scheme.
+
+   This scheme was added in Python 1.5.2b2; before that time, only PATCHLEVEL
+   was available.  To test for presence of the scheme, test for
+   defined(PY_MAJOR_VERSION).
+
+   When the major or minor version changes, the VERSION variable in
+   configure.in must also be changed.
+
+   There is also (independent) API version information in modsupport.h.
+*/
+
+/* Values for PY_RELEASE_LEVEL */
+#define PY_RELEASE_LEVEL_ALPHA	0xA
+#define PY_RELEASE_LEVEL_BETA	0xB
+#define PY_RELEASE_LEVEL_GAMMA	0xC     /* For release candidates */
+#define PY_RELEASE_LEVEL_FINAL	0xF	/* Serial should be 0 here */
+					/* Higher for patch releases */
+
+/* Version parsed out into numeric values */
+#define PY_MAJOR_VERSION	2
+#define PY_MINOR_VERSION	7
+#define PY_MICRO_VERSION	13
+#define PY_RELEASE_LEVEL	PY_RELEASE_LEVEL_FINAL
+#define PY_RELEASE_SERIAL	0
+
+/* Version as a string */
+#define PY_VERSION		"2.7.13"
+
+/* PyPy version as a string */
+#define PYPY_VERSION "5.10.0"
+#define PYPY_VERSION_NUM  0x050A0000
+
+/* Defined to mean a PyPy where cpyext holds more regular references
+   to PyObjects, e.g. staying alive as long as the internal PyPy object
+   stays alive. */
+#define PYPY_CPYEXT_GC      1
+#define PyPy_Borrow(a, b)   ((void) 0)
+
+/* Subversion Revision number of this file (not of the repository).
+ * Empty since Mercurial migration. */
+#define PY_PATCHLEVEL_REVISION  ""
+
+/* Version as a single 4-byte hex number, e.g. 0x010502B2 == 1.5.2b2.
+   Use this for numeric comparisons, e.g. #if PY_VERSION_HEX >= ... */
+#define PY_VERSION_HEX ((PY_MAJOR_VERSION << 24) | \
+			(PY_MINOR_VERSION << 16) | \
+			(PY_MICRO_VERSION <<  8) | \
+			(PY_RELEASE_LEVEL <<  4) | \
+			(PY_RELEASE_SERIAL << 0))

### 2018-03-22 test_env/pypy-withpatch/pypy/module/cpyext/test/test_pyerrors.py

@@ -1,0 +1,439 @@
+import pytest
+import sys
+import StringIO
+
+from pypy.module.cpyext.state import State
+from pypy.module.cpyext.test.test_api import BaseApiTest
+from pypy.module.cpyext.test.test_cpyext import AppTestCpythonExtensionBase
+from rpython.rtyper.lltypesystem import rffi
+
+class TestExceptions(BaseApiTest):
+    def test_GivenExceptionMatches(self, space, api):
+        old_style_exception = space.appexec([], """():
+            class OldStyle:
+                pass
+            return OldStyle
+        """)
+        exc_matches = api.PyErr_GivenExceptionMatches
+
+        string_exception = space.wrap('exception')
+        instance = space.call_function(space.w_ValueError)
+        old_style_instance = space.call_function(old_style_exception)
+        assert exc_matches(string_exception, string_exception)
+        assert exc_matches(old_style_exception, old_style_exception)
+        assert not exc_matches(old_style_exception, space.w_Exception)
+        assert exc_matches(instance, space.w_ValueError)
+        assert exc_matches(old_style_instance, old_style_exception)
+        assert exc_matches(space.w_ValueError, space.w_ValueError)
+        assert exc_matches(space.w_IndexError, space.w_LookupError)
+        assert not exc_matches(space.w_ValueError, space.w_LookupError)
+
+        exceptions = space.newtuple([space.w_LookupError, space.w_ValueError])
+        assert exc_matches(space.w_ValueError, exceptions)
+
+    def test_ExceptionMatches(self, space, api):
+        api.PyErr_SetObject(space.w_ValueError, space.wrap("message"))
+        assert api.PyErr_ExceptionMatches(space.w_Exception)
+        assert api.PyErr_ExceptionMatches(space.w_ValueError)
+        assert not api.PyErr_ExceptionMatches(space.w_TypeError)
+
+        api.PyErr_Clear()
+
+    def test_Occurred(self, space, api):
+        assert not api.PyErr_Occurred()
+        string = rffi.str2charp("spam and eggs")
+        api.PyErr_SetString(space.w_ValueError, string)
+        rffi.free_charp(string)
+        assert api.PyErr_Occurred() is space.w_ValueError
+
+        api.PyErr_Clear()
+
+    def test_SetObject(self, space, api):
+        api.PyErr_SetObject(space.w_ValueError, space.wrap("a value"))
+        assert api.PyErr_Occurred() is space.w_ValueError
+        state = space.fromcache(State)
[…]

### 2018-03-22 test_env/pypy-withpatch/pypy/module/cpyext/test/test_sequence.py

@@ -1,0 +1,296 @@
+from rpython.rtyper.lltypesystem import rffi
+from pypy.interpreter.error import OperationError
+from pypy.module.cpyext.test.test_api import BaseApiTest, raises_w
+from pypy.module.cpyext.test.test_cpyext import AppTestCpythonExtensionBase
+from pypy.module.cpyext.sequence import (
+    PySequence_Fast, PySequence_Contains, PySequence_Index,
+    PySequence_GetItem, PySequence_SetItem, PySequence_DelItem)
+
+import pytest
+
+class TestSequence(BaseApiTest):
+    def test_check(self, space, api):
+        assert api.PySequence_Check(space.newlist([]))
+        assert not api.PySequence_Check(space.newdict())
+
+    def test_sequence_api(self, space, api):
+        w_l = space.wrap([1, 2, 3, 4])
+        assert api.PySequence_Fast(w_l, "message") is w_l
+
+        assert space.int_w(api.PySequence_Fast_GET_ITEM(w_l, 1)) == 2
+        assert api.PySequence_Fast_GET_SIZE(w_l) == 4
+
+        w_set = space.wrap(set((1, 2, 3, 4)))
+        w_seq = api.PySequence_Fast(w_set, "message")
+        assert space.type(w_seq) is space.w_list
+        assert space.len_w(w_seq) == 4
+
+        w_seq = api.PySequence_Tuple(w_set)
+        assert space.type(w_seq) is space.w_tuple
+        assert sorted(space.unwrap(w_seq)) == [1, 2, 3, 4]
+
+        w_seq = api.PySequence_List(w_set)
+        assert space.type(w_seq) is space.w_list
+        assert sorted(space.unwrap(w_seq)) == [1, 2, 3, 4]
+
+    def test_repeat(self, space, api):
+        def test(seq, count):
+            w_seq = space.wrap(seq)
+            w_repeated = api.PySequence_Repeat(w_seq, count)
+            assert space.eq_w(w_repeated, space.wrap(seq * count))
+
+        test((1, 2, 3, 4), 3)
+        test([1, 2, 3, 4], 3)
+
+    def test_concat(self, space, api):
+        w_t1 = space.wrap(range(4))
+        w_t2 = space.wrap(range(4, 8))
+        assert space.unwrap(api.PySequence_Concat(w_t1, w_t2)) == range(8)
+
+    def test_inplace_concat(self, space, api):
[…]

### 2018-03-22 test_env/pypy-withpatch/pypy/module/cpyext/test/test_unicodeobject.py

@@ -1,0 +1,654 @@
+# encoding: utf-8
+import pytest
+from pypy.module.cpyext.test.test_api import BaseApiTest, raises_w
+from pypy.module.cpyext.test.test_cpyext import AppTestCpythonExtensionBase
+from pypy.module.cpyext.unicodeobject import (
+    Py_UNICODE, PyUnicodeObject, new_empty_unicode)
+from pypy.module.cpyext.api import PyObjectP, PyObject
+from pypy.module.cpyext.pyobject import Py_DecRef, from_ref
+from rpython.rtyper.lltypesystem import rffi, lltype
+import sys, py
+from pypy.module.cpyext.unicodeobject import *
+
+class AppTestUnicodeObject(AppTestCpythonExtensionBase):
+    def test_unicodeobject(self):
+        module = self.import_extension('foo', [
+            ("get_hello1", "METH_NOARGS",
+             """
+                 return PyUnicode_FromStringAndSize(
+                     "Hello world<should not be included>", 11);
+             """),
+            ("test_GetSize", "METH_NOARGS",
+             """
+                 PyObject* s = PyUnicode_FromString("Hello world");
+                 int result = 0;
+
+                 if(PyUnicode_GetSize(s) != 11) {
+                     result = -PyUnicode_GetSize(s);
+                 }
+#ifdef PYPY_VERSION
+                 // Slightly silly test that tp_basicsize is reasonable.
+                 if(s->ob_type->tp_basicsize != sizeof(void*)*7)
+                     result = s->ob_type->tp_basicsize;
+#endif  // PYPY_VERSION
+                 Py_DECREF(s);
+                 return PyLong_FromLong(result);
+             """),
+            ("test_GetSize_exception", "METH_NOARGS",
+             """
+                 PyObject* f = PyFloat_FromDouble(1.0);
+                 PyUnicode_GetSize(f);
+
+                 Py_DECREF(f);
+                 return NULL;
+             """),
+             ("test_is_unicode", "METH_VARARGS",
+             """
+                return PyBool_FromLong(PyUnicode_Check(PyTuple_GetItem(args, 0)));
+             """)])
+        assert module.get_hello1() == u'Hello world'
+        assert module.test_GetSize() == 0
+        raises(TypeError, module.test_GetSize_exception)
+
+        assert module.test_is_unicode(u"")
[…]

### 2018-03-22 test_env/pypy-withpatch/pypy/module/cpyext/unicodeobject.py

@@ -1,0 +1,760 @@
+from pypy.interpreter.error import OperationError, oefmt
+from rpython.rtyper.lltypesystem import rffi, lltype
+from pypy.module.unicodedata import unicodedb
+from pypy.module.cpyext.api import (
+    CANNOT_FAIL, Py_ssize_t, build_type_checkers_flags, cpython_api,
+    bootstrap_function, CONST_STRING,
+    CONST_WSTRING, slot_function, cts, parse_dir)
+from pypy.module.cpyext.pyerrors import PyErr_BadArgument
+from pypy.module.cpyext.pyobject import (
+    PyObject, PyObjectP, Py_DecRef, make_ref, from_ref, track_reference,
+    make_typedescr, get_typedescr)
+from pypy.module.cpyext.bytesobject import PyString_Check
+from pypy.module.sys.interp_encoding import setdefaultencoding
+from pypy.module._codecs.interp_codecs import CodecState
+from pypy.interpreter import unicodehelper
+from pypy.objspace.std import unicodeobject
+from rpython.rlib import rstring, runicode
+from rpython.tool.sourcetools import func_renamer
+import sys
+
+## See comment in bytesobject.py.
+
+cts.parse_header(parse_dir / 'cpyext_unicodeobject.h')
+PyUnicodeObject = cts.gettype('PyUnicodeObject*')
+Py_UNICODE = cts.gettype('Py_UNICODE')
+
+@bootstrap_function
+def init_unicodeobject(space):
+    make_typedescr(space.w_unicode.layout.typedef,
+                   basestruct=PyUnicodeObject.TO,
+                   attach=unicode_attach,
+                   dealloc=unicode_dealloc,
+                   realize=unicode_realize)
+
+# Buffer for the default encoding (used by PyUnicde_GetDefaultEncoding)
+DEFAULT_ENCODING_SIZE = 100
+default_encoding = lltype.malloc(rffi.CCHARP.TO, DEFAULT_ENCODING_SIZE,
+                                 flavor='raw', zero=True)
+
+PyUnicode_Check, PyUnicode_CheckExact = build_type_checkers_flags("Unicode")
+
+
+def new_empty_unicode(space, length):
+    """
+    Allocate a PyUnicodeObject and its buffer, but without a corresponding
[…]

### 2018-03-22 test_env/pypy-withpatch/pypy/module/test_lib_pypy/cffi_tests/cffi1/test_recompiler.py

@@ -1,0 +1,2290 @@
+# Generated by pypy/tool/import_cffi.py
+
+import sys, os, py
+from cffi import FFI, VerificationError, FFIError, CDefError
+from cffi import recompiler
+from pypy.module.test_lib_pypy.cffi_tests.udir import udir
+from pypy.module.test_lib_pypy.cffi_tests.support import u, long
+from pypy.module.test_lib_pypy.cffi_tests.support import FdWriteCapture, StdErrCapture
+
+try:
+    import importlib
+except ImportError:
+    importlib = None
+
+
+def check_type_table(input, expected_output, included=None):
+    ffi = FFI()
+    if included:
+        ffi1 = FFI()
+        ffi1.cdef(included)
+        ffi.include(ffi1)
+    ffi.cdef(input)
+    recomp = recompiler.Recompiler(ffi, 'testmod')
+    recomp.collect_type_table()
+    assert ''.join(map(str, recomp.cffi_types)) == expected_output
+
+def verify(ffi, module_name, source, *args, **kwds):
+    no_cpp = kwds.pop('no_cpp', False)
+    kwds.setdefault('undef_macros', ['NDEBUG'])
+    module_name = '_CFFI_' + module_name
+    ffi.set_source(module_name, source)
+    if not os.environ.get('NO_CPP') and not no_cpp:   # test the .cpp mode too
+        kwds.setdefault('source_extension', '.cpp')
+        source = 'extern "C" {\n%s\n}' % (source,)
+    elif sys.platform != 'win32':
+        # add '-Werror' to the existing 'extra_compile_args' flags
+        kwds['extra_compile_args'] = (kwds.get('extra_compile_args', []) +
+                                      ['-Werror'])
+    return recompiler._verify(ffi, module_name, source, *args, **kwds)
+
+def test_set_source_no_slashes():
+    ffi = FFI()
+    py.test.raises(ValueError, ffi.set_source, "abc/def", None)
+    py.test.raises(ValueError, ffi.set_source, "abc/def", "C code")
+
+
+def test_type_table_func():
+    check_type_table("double sin(double);",
+                     "(FUNCTION 1)(PRIMITIVE 14)(FUNCTION_END 0)")
+    check_type_table("float sin(double);",
[…]

### 2018-03-22 test_env/pypy-withpatch/pypy/module/time/interp_time.py

@@ -1,0 +1,656 @@
+from rpython.rtyper.tool import rffi_platform as platform
+from rpython.rtyper.lltypesystem import rffi
+from pypy.interpreter.error import OperationError, oefmt
+from pypy.interpreter.gateway import unwrap_spec
+from rpython.rtyper.lltypesystem import lltype
+from rpython.rlib.rarithmetic import intmask
+from rpython.rlib import rposix, rtime
+from rpython.translator.tool.cbuild import ExternalCompilationInfo
+import os
+import sys
+import time as pytime
+
+_POSIX = os.name == "posix"
+_WIN = os.name == "nt"
+_CYGWIN = sys.platform == "cygwin"
+
+_time_zones = []
+if _CYGWIN:
+    _time_zones = ["GMT-12", "GMT-11", "GMT-10", "GMT-9", "GMT-8", "GMT-7",
+                   "GMT-6", "GMT-5", "GMT-4", "GMT-3", "GMT-2", "GMT-1",
+                   "GMT",  "GMT+1", "GMT+2", "GMT+3", "GMT+4", "GMT+5",
+                   "GMT+6",  "GMT+7", "GMT+8", "GMT+9", "GMT+10", "GMT+11",
+                   "GMT+12",  "GMT+13", "GMT+14"]
+
+if _WIN:
+    # Interruptible sleeps on Windows:
+    # We install a specific Console Ctrl Handler which sets an 'event'.
+    # time.sleep() will actually call WaitForSingleObject with the desired
+    # timeout.  On Ctrl-C, the signal handler is called, the event is set,
+    # and the wait function exits.
+    from rpython.rlib import rwin32
+    from pypy.interpreter.error import wrap_windowserror, wrap_oserror
+    from rpython.rlib import rthread as thread
+
+    eci = ExternalCompilationInfo(
+        includes = ['windows.h'],
+        post_include_bits = [
+            "RPY_EXTERN\n"
+            "BOOL pypy_timemodule_setCtrlHandler(HANDLE event);"],
+        separate_module_sources=['''
+            static HANDLE interrupt_event;
+
+            static BOOL WINAPI CtrlHandlerRoutine(
+              DWORD dwCtrlType)
+            {
+                SetEvent(interrupt_event);
+                /* allow other default handlers to be called.
+                 * Default Python handler will setup the
+                 * KeyboardInterrupt exception.
+                 */
+                return 0;
+            }
+
[…]

### 2018-03-22 test_env/pypy-withpatch/pypy/tool/release/repackage.sh

@@ -1,0 +1,90 @@
+# Edit these appropriately before running this script
+pmaj=2  # python main version
+pmin=7  # python minor version
+maj=5
+min=8
+rev=0
+branchname=release-pypy$pmaj.$pmin-$maj.x # ==OR== release-$maj.x  # ==OR== release-$maj.$min.x
+tagname=release-pypy$pmaj.$pmin-v$maj.$min.$rev  # ==OR== release-$maj.$min
+
+echo checking hg log -r $branchname
+hg log -r $branchname || exit 1
+echo checking hg log -r $tagname
+hg log -r $tagname || exit 1
+hgrev=`hg id -r $tagname -i`
+
+rel=pypy$pmaj-v$maj.$min.$rev
+# The script should be run in an empty in the pypy tree, i.e. pypy/tmp
+if [ "`ls . | wc -l`" != "0" ]
+then
+    echo this script must be run in an empty directory
+    exit -1
+fi
+
+# Download latest builds from the buildmaster, rename the top
+# level directory, and repackage ready to be uploaded to bitbucket
+for plat in linux linux64 linux-armhf-raspbian linux-armel osx64 s390x
+  do
+    echo downloading package for $plat
+    if wget -q --show-progress http://buildbot.pypy.org/nightly/$branchname/pypy-c-jit-latest-$plat.tar.bz2
+    then
+        echo $plat downloaded 
+    else
+        echo $plat no download available
+        continue
+    fi
+    hgcheck=`tar -tf pypy-c-jit-latest-$plat.tar.bz2 |head -n1 | cut -d- -f5`
+    if [ "$hgcheck" != "$hgrev" ]
+    then
+        echo xxxxxxxxxxxxxxxxxxxxxx
+        echo $plat hg tag mismatch, expected $hgrev, got $hgcheck
+        echo xxxxxxxxxxxxxxxxxxxxxx
+        rm pypy-c-jit-latest-$plat.tar.bz2
+        continue
+    fi
+    tar -xf pypy-c-jit-latest-$plat.tar.bz2
+    rm pypy-c-jit-latest-$plat.tar.bz2
+    plat_final=$plat
+    if [ $plat = linux ]; then
+        plat_final=linux32
+    fi
+    mv pypy-c-jit-*-$plat $rel-$plat_final
+    echo packaging $plat_final
[…]

### 2018-03-22 test_env/pypy-withpatch/rpython/annotator/bookkeeper.py

@@ -1,0 +1,613 @@
+"""
+The Bookkeeper class.
+"""
+
+from __future__ import absolute_import
+
+import sys, types, inspect, weakref
+from contextlib import contextmanager
+from collections import OrderedDict
+
+from rpython.flowspace.model import Constant
+from rpython.flowspace.bytecode import cpython_code_signature
+from rpython.annotator.model import (
+    SomeOrderedDict, SomeString, SomeChar, SomeFloat, unionof, SomeInstance,
+    SomeDict, SomeBuiltin, SomePBC, SomeInteger, TLS, SomeUnicodeCodePoint,
+    s_None, s_ImpossibleValue, SomeBool, SomeTuple, SomeException,
+    SomeImpossibleValue, SomeUnicodeString, SomeList, HarmlesslyBlocked,
+    SomeWeakRef, SomeByteArray, SomeConstantType, SomeProperty,
+    AnnotatorError)
+from rpython.annotator.classdesc import ClassDef, ClassDesc
+from rpython.annotator.listdef import ListDef, ListItem
+from rpython.annotator.dictdef import DictDef
+from rpython.annotator import description
+from rpython.annotator.signature import annotationoftype
+from rpython.annotator.argument import simple_args
+from rpython.annotator.specialize import memo
+from rpython.rlib.objectmodel import r_dict, r_ordereddict, Symbolic
+from rpython.tool.algo.unionfind import UnionFind
+from rpython.rtyper import extregistry
+
+
+BUILTIN_ANALYZERS = {}
+
+def analyzer_for(func):
+    def wrapped(ann_func):
+        BUILTIN_ANALYZERS[func] = ann_func
+        return ann_func
+    return wrapped
+
+class Bookkeeper(object):
+    """The log of choices that have been made while analysing the operations.
+    It ensures that the same 'choice objects' will be returned if we ask
+    again during reflowing.  Like ExecutionContext, there is an implicit
+    Bookkeeper that can be obtained from a thread-local variable.
+
+    Currently used for factories and user-defined classes."""
+
+    def __setstate__(self, dic):
+        self.__dict__.update(dic) # normal action
+        self.register_builtins()
+
+    def __init__(self, annotator):
+        self.annotator = annotator
+        self.policy = annotator.policy
+        self.descs = {}          # map Python objects to their XxxDesc wrappers
+        self.methoddescs = {}    # map (funcdesc, classdef) to the MethodDesc
[…]

### 2018-03-22 test_env/pypy-withpatch/rpython/jit/metainterp/optimizeopt/optimizer.py

@@ -1,0 +1,872 @@
+from rpython.jit.metainterp import jitprof, resume, compile
+from rpython.jit.metainterp.executor import execute_nonspec_const
+from rpython.jit.metainterp.history import Const, ConstInt, ConstPtr
+from rpython.jit.metainterp.optimizeopt.intutils import IntBound,\
+     ConstIntBound, MININT, MAXINT, IntUnbounded
+from rpython.jit.metainterp.optimizeopt.util import make_dispatcher_method
+from rpython.jit.metainterp.resoperation import rop, AbstractResOp, GuardResOp,\
+     OpHelpers
+from rpython.jit.metainterp.optimizeopt import info
+from rpython.jit.metainterp.optimize import InvalidLoop
+from rpython.jit.metainterp.typesystem import llhelper
+from rpython.rlib.objectmodel import specialize, we_are_translated
+from rpython.rlib.debug import debug_print
+from rpython.rtyper import rclass
+from rpython.rtyper.lltypesystem import llmemory
+from rpython.jit.metainterp.optimize import SpeculativeError
+
+
+
+
+CONST_0      = ConstInt(0)
+CONST_1      = ConstInt(1)
+CONST_ZERO_FLOAT = Const._new(0.0)
+llhelper.CONST_NULLREF = llhelper.CONST_NULL
+REMOVED = AbstractResOp()
+
+class LoopInfo(object):
+    label_op = None
+
+class BasicLoopInfo(LoopInfo):
+    def __init__(self, inputargs, quasi_immutable_deps, jump_op):
+        self.inputargs = inputargs
+        self.jump_op = jump_op
+        self.quasi_immutable_deps = quasi_immutable_deps
+        self.extra_same_as = []
+        self.extra_before_label = []
+
+    def final(self):
+        return True
+
+    def post_loop_compilation(self, loop, jitdriver_sd, metainterp, jitcell_token):
+        pass
+
+
+class OptimizationResult(object):
+    def __init__(self, opt, op):
+        self.opt = opt
+        self.op = op
+
+    def callback(self):
+        self.opt.propagate_postprocess(self.op)
+
+
+class Optimization(object):
+    next_optimization = None
+    potential_extra_ops = None
+
+    def __init__(self):
+        pass # make rpython happy
+
+    def send_extra_operation(self, op, opt=None):
+        self.optimizer.send_extra_operation(op, opt)
+
+    def propagate_forward(self, op):
+        raise NotImplementedError
+
[…]

### 2018-03-22 test_env/pypy-withpatch/rpython/memory/gc/minimark.py

@@ -1,0 +1,2150 @@
+""" MiniMark GC.
+
+Environment variables can be used to fine-tune the following parameters:
+
+ PYPY_GC_NURSERY         The nursery size.  Defaults to 1/2 of your cache or
+                         '4M'.  Small values
+                         (like 1 or 1KB) are useful for debugging.
+
+ PYPY_GC_NURSERY_CLEANUP The interval at which nursery is cleaned up. Must
+                         be smaller than the nursery size and bigger than the
+                         biggest object we can allotate in the nursery.
+
+ PYPY_GC_MAJOR_COLLECT   Major collection memory factor.  Default is '1.82',
+                         which means trigger a major collection when the
+                         memory consumed equals 1.82 times the memory
+                         really used at the end of the previous major
+                         collection.
+
+ PYPY_GC_GROWTH          Major collection threshold's max growth rate.
+                         Default is '1.4'.  Useful to collect more often
+                         than normally on sudden memory growth, e.g. when
+                         there is a temporary peak in memory usage.
+
+ PYPY_GC_MAX             The max heap size.  If coming near this limit, it
+                         will first collect more often, then raise an
+                         RPython MemoryError, and if that is not enough,
+                         crash the program with a fatal error.  Try values
+                         like '1.6GB'.
+
+ PYPY_GC_MAX_DELTA       The major collection threshold will never be set
+                         to more than PYPY_GC_MAX_DELTA the amount really
+                         used after a collection.  Defaults to 1/8th of the
+                         total RAM size (which is constrained to be at most
+                         2/3/4GB on 32-bit systems).  Try values like '200MB'.
+
+ PYPY_GC_MIN             Don't collect while the memory size is below this
+                         limit.  Useful to avoid spending all the time in
+                         the GC in very small programs.  Defaults to 8
+                         times the nursery.
+
+ PYPY_GC_DEBUG           Enable extra checks around collections that are
+                         too slow for normal use.  Values are 0 (off),
+                         1 (on major collections) or 2 (also on minor
+                         collections).
+"""
+# XXX Should find a way to bound the major collection threshold by the
[…]

### 2018-03-22 test_env/pypy-withpatch/rpython/rlib/debug.py

@@ -1,0 +1,534 @@
+import sys
+import time
+
+from rpython.rtyper.extregistry import ExtRegistryEntry
+from rpython.rlib.objectmodel import we_are_translated
+from rpython.rlib.rarithmetic import is_valid_int
+from rpython.rtyper.extfunc import register_external
+from rpython.rtyper.lltypesystem import lltype
+from rpython.rtyper.lltypesystem import rffi
+from rpython.translator.tool.cbuild import ExternalCompilationInfo
+
+# Expose these here (public interface)
+from rpython.rtyper.debug import (
+    ll_assert, FatalError, fatalerror, fatalerror_notb, debug_print_traceback,
+    ll_assert_not_none)
+
+
+class DebugLog(list):
+    def debug_print(self, *args):
+        self.append(('debug_print',) + args)
+
+    def debug_start(self, category, time=None):
+        self.append(('debug_start', category, time))
+
+    def debug_stop(self, category, time=None):
+        for i in xrange(len(self) - 1, -1, -1):
+            if self[i][0] == 'debug_start':
+                assert self[i][1] == category, (
+                    "nesting error: starts with %r but stops with %r" %
+                    (self[i][1], category))
+                starttime = self[i][2]
+                if starttime is not None or time is not None:
+                    self[i:] = [(category, starttime, time, self[i + 1:])]
+                else:
+                    self[i:] = [(category, self[i + 1:])]
+                return
+        assert False, ("nesting error: no start corresponding to stop %r" %
+                       (category,))
+
+    def __repr__(self):
+        import pprint
+        return pprint.pformat(list(self))
+
+_log = None       # patched from tests to be an object of class DebugLog
+                  # or compatible
+
+def debug_print(*args):
+    for arg in args:
+        print >> sys.stderr, arg,
+    print >> sys.stderr
+    if _log is not None:
+        _log.debug_print(*args)
+
+class Entry(ExtRegistryEntry):
+    _about_ = debug_print
+
+    def compute_result_annotation(self, *args_s):
+        return None
+
+    def specialize_call(self, hop):
+        vlist = hop.inputargs(*hop.args_r)
+        hop.exception_cannot_occur()
+        t = hop.rtyper.annotator.translator
+        if t.config.translation.log:
+            hop.genop('debug_print', vlist)
+
+
+if sys.stderr.isatty():
+    _start_colors_1 = "\033[1m\033[31m"
[…]

### 2018-03-22 test_env/pypy-withpatch/rpython/rlib/test/test_objectmodel.py

@@ -1,0 +1,850 @@
+from collections import OrderedDict
+import py
+from rpython.rlib.objectmodel import (
+    r_dict, UnboxedValue, Symbolic, compute_hash, compute_identity_hash,
+    compute_unique_id, current_object_addr_as_int, we_are_translated,
+    prepare_dict_update, reversed_dict, specialize, enforceargs, newlist_hint,
+    resizelist_hint, is_annotation_constant, always_inline, NOT_CONSTANT,
+    iterkeys_with_hash, iteritems_with_hash, contains_with_hash,
+    setitem_with_hash, getitem_with_hash, delitem_with_hash, import_from_mixin,
+    fetch_translated_config, try_inline, delitem_if_value_is, move_to_end)
+from rpython.translator.translator import TranslationContext, graphof
+from rpython.rtyper.test.tool import BaseRtypingTest
+from rpython.rtyper.test.test_llinterp import interpret
+from rpython.conftest import option
+
+def strange_key_eq(key1, key2):
+    return key1[0] == key2[0]   # only the 1st character is relevant
+def strange_key_hash(key):
+    return ord(key[0])
+
+def play_with_r_dict(d):
+    d['hello'] = 41
+    d['hello'] = 42
+    assert d['hi there'] == 42
+    try:
+        unexpected = d["dumb"]
+    except KeyError:
+        pass
+    else:
+        assert False, "should have raised, got %s" % unexpected
+    assert len(d) == 1
+    assert 'oops' not in d
+
+    count = 0
+    for x in d:
+        assert x == 'hello'
+        count += 1
+    assert count == 1
+
+    assert d.get('hola', -1) == 42
+    assert d.get('salut', -1) == -1
+    d1 = d.copy()
+    del d['hu!']
+    assert len(d) == 0
+    assert d1.keys() == ['hello']
+    d.update(d1)
+    assert d.values() == [42]
+    lst = d.items()
+    assert len(lst) == 1 and len(lst[0]) == 2
+    assert lst[0][0] == 'hello' and lst[0][1] == 42
+
+    count = 0
+    for x in d.iterkeys():
+        assert x == 'hello'
+        count += 1
+    assert count == 1
+
+    count = 0
+    for x in d.itervalues():
+        assert x == 42
+        count += 1
+    assert count == 1
+
+    count = 0
+    for x in d.iteritems():
+        assert len(x) == 2 and x[0] == 'hello' and x[1] == 42
+        count += 1
[…]

### 2018-03-22 test_env/pypy-withpatch/rpython/rlib/test/test_rwin32.py

@@ -1,0 +1,70 @@
+# encoding: utf-8
+import os, py
+if os.name != 'nt':
+    py.test.skip('tests for win32 only')
+
+from rpython.rlib import rwin32
+from rpython.tool.udir import udir
+
+
+def test_get_osfhandle():
+    fid = open(str(udir.join('validate_test.txt')), 'w')
+    fd = fid.fileno()
+    rwin32.get_osfhandle(fd)
+    fid.close()
+    py.test.raises(OSError, rwin32.get_osfhandle, fd)
+    rwin32.get_osfhandle(0)
+
+def test_get_osfhandle_raising():
+    #try to test what kind of exception get_osfhandle raises w/out fd validation
+    py.test.skip('Crashes python')
+    fid = open(str(udir.join('validate_test.txt')), 'w')
+    fd = fid.fileno()
+    fid.close()
+    def validate_fd(fd):
+        return 1
+    _validate_fd = rwin32.validate_fd
+    rwin32.validate_fd = validate_fd
+    raises(WindowsError, rwin32.get_osfhandle, fd)
+    rwin32.validate_fd = _validate_fd
+
+def test_open_process():
+    pid = rwin32.GetCurrentProcessId()
+    assert pid != 0
+    handle = rwin32.OpenProcess(rwin32.PROCESS_QUERY_INFORMATION, False, pid)
+    rwin32.CloseHandle(handle)
+    py.test.raises(WindowsError, rwin32.OpenProcess, rwin32.PROCESS_TERMINATE, False, 0)
+
+def test_terminate_process():
+    import subprocess, signal, sys
+    proc = subprocess.Popen([sys.executable, "-c",
+                         "import time;"
+                         "time.sleep(10)",
+                         ],
+                        ) 
+    print proc.pid
+    handle = rwin32.OpenProcess(rwin32.PROCESS_ALL_ACCESS, False, proc.pid)
+    assert rwin32.TerminateProcess(handle, signal.SIGTERM) == 1
+    rwin32.CloseHandle(handle)
+    assert proc.wait() == signal.SIGTERM
+ 
+@py.test.mark.dont_track_allocations('putenv intentionally keeps strings alive')
+def test_wenviron():
+    name, value = u'PYPY_TEST_日本', u'foobar日本'
+    rwin32._wputenv(name, value)
+    assert rwin32._wgetenv(name) == value
+    env = dict(rwin32._wenviron_items())
+    assert env[name] == value
+    for key, value in env.iteritems():
[…]

### 2018-03-22 test_env/pypy-withpatch/rpython/translator/c/test/test_standalone.py

@@ -1,0 +1,1442 @@
+import py
+import sys, os, re
+
+from rpython.config.translationoption import get_combined_translation_config
+from rpython.config.translationoption import SUPPORT__THREAD
+from rpython.rlib.objectmodel import keepalive_until_here
+from rpython.rlib.rarithmetic import r_longlong
+from rpython.rlib.debug import ll_assert, have_debug_prints, debug_flush
+from rpython.rlib.debug import debug_print, debug_start, debug_stop
+from rpython.rlib.debug import debug_offset, have_debug_prints_for
+from rpython.rlib.entrypoint import entrypoint_highlevel, secondary_entrypoints
+from rpython.rtyper.lltypesystem import lltype
+from rpython.translator.translator import TranslationContext
+from rpython.translator.backendopt import all
+from rpython.translator.c.genc import CStandaloneBuilder, ExternalCompilationInfo
+from rpython.annotator.listdef import s_list_of_strings
+from rpython.tool.udir import udir
+from rpython.translator import cdir
+from rpython.conftest import option
+from rpython.rlib.jit import JitDriver
+
+def setup_module(module):
+    if os.name == 'nt':
+        # Do not open dreaded dialog box on segfault
+        import ctypes
+        SEM_NOGPFAULTERRORBOX = 0x0002 # From MSDN
+        if hasattr(ctypes.windll.kernel32, 'GetErrorMode'):
+            old_err_mode = ctypes.windll.kernel32.GetErrorMode()
+            new_err_mode = old_err_mode | SEM_NOGPFAULTERRORBOX
+            ctypes.windll.kernel32.SetErrorMode(new_err_mode)
+            module.old_err_mode = old_err_mode
+
+def teardown_module(module):
+    if os.name == 'nt' and hasattr(module, 'old_err_mode'):
+        import ctypes
+        ctypes.windll.kernel32.SetErrorMode(module.old_err_mode)
+
+class StandaloneTests(object):
+    config = None
+
+    def compile(self, entry_point, debug=True, shared=False,
+                stackcheck=False, entrypoints=None, local_icon=None):
+        t = TranslationContext(self.config)
+        ann = t.buildannotator()
+        ann.build_types(entry_point, [s_list_of_strings])
+        if entrypoints is not None:
+            anns = {}
+            for func, annotation in secondary_entrypoints['test']:
+                anns[func] = annotation
+            for item in entrypoints:
+                ann.build_types(item, anns[item])
[…]

### 2018-03-22 test_env/pypy-withpatch/rpython/translator/platform/__init__.py

@@ -1,0 +1,352 @@
+"""Platform-specific support for compiling/executing C sources."""
+
+import py, os, sys
+
+from rpython.tool.runsubprocess import run_subprocess as _run_subprocess
+from rpython.tool.udir import udir
+from rpython.tool.version import rpythonroot
+from rpython.tool.ansi_print import AnsiLogger
+
+log = AnsiLogger("platform")
+
+
+class CompilationError(Exception):
+    def __init__(self, out, err):
+        self.out = out.replace('\r\n', '\n')
+        self.err = err.replace('\r\n', '\n')
+
+    def __repr__(self):
+        if self.err:
+            attr = 'err'
+        else:
+            attr = 'out'
+        text = getattr(self, attr).replace('\n', '\n\t')
+        return 'CompilationError(%s="""\n\t%s""")' % (attr, text)
+
+    __str__ = __repr__
+
+class ExecutionResult(object):
+    def __init__(self, returncode, out, err):
+        self.returncode = returncode
+        self.out = out.replace('\r\n', '\n')
+        self.err = err.replace('\r\n', '\n')
+
+    def __repr__(self):
+        return "<ExecutionResult retcode=%d>" % (self.returncode,)
+
+class Platform(object):
+    name = "abstract platform"
+    c_environ = None
+
+    relevant_environ = ()
+    log_errors = True
+
+    so_prefixes = ('',)
+
+    extra_libs = ()
+
+    def __init__(self, cc):
+        if self.__class__ is Platform:
+            raise TypeError("You should not instantiate Platform class directly")
+        self.cc = cc
+
+    def compile(self, cfiles, eci, outputfilename=None, standalone=True):
+        ofiles = self._compile_o_files(cfiles, eci, standalone)
+        return self._finish_linking(ofiles, eci, outputfilename, standalone)
+
+    def _all_cfiles(self, cfiles, eci):
+        seen = set()
+        result = []
+        for cfile in list(cfiles) + list(eci.separate_module_files):
+            cfile = py.path.local(cfile)
+            if cfile not in seen:
+                seen.add(cfile)
+                result.append(cfile)
+        return result
+
+    def _compile_o_files(self, cfiles, eci, standalone=True):
+        cfiles = self._all_cfiles(cfiles, eci)
+        compile_args = self._compile_args_from_eci(eci, standalone)
+        ofiles = []
+        for cfile in cfiles:
+            # Windows hack: use masm for files ending in .asm
+            if str(cfile).lower().endswith('.asm'):
[…]

### 2018-03-22 test_env/pypy-withpatch/rpython/translator/platform/windows.py

@@ -1,0 +1,591 @@
+"""Support for Windows."""
+
+import py, os, sys, re, shutil
+
+from rpython.translator.platform import CompilationError
+from rpython.translator.platform import log, _run_subprocess
+from rpython.translator.platform import Platform, posix
+
+import rpython
+rpydir = str(py.path.local(rpython.__file__).join('..'))
+
+def _get_compiler_type(cc, x64_flag):
+    if not cc:
+        cc = os.environ.get('CC','')
+    if not cc:
+        return MsvcPlatform(x64=x64_flag)
+    elif cc.startswith('mingw') or cc == 'gcc':
+        return MingwPlatform(cc)
+    return MsvcPlatform(cc=cc, x64=x64_flag)
+
+def Windows(cc=None):
+    return _get_compiler_type(cc, False)
+
+def Windows_x64(cc=None):
+    raise Exception("Win64 is not supported.  You must either build for Win32"
+                    " or contribute the missing support in PyPy.")
+    return _get_compiler_type(cc, True)
+
+def _find_vcvarsall(version):
+    # copied from setuptools.msvc9_support.py
+    from distutils.msvc9compiler import Reg
+    VC_BASE = r'Software\%sMicrosoft\DevDiv\VCForPython\%0.1f'
+    key = VC_BASE % ('', version)
+    try:
+        # Per-user installs register the compiler path here
+        productdir = Reg.get_value(key, "installdir")
+    except KeyError:
+        try:
+            # All-user installs on a 64-bit system register here
+            key = VC_BASE % ('Wow6432Node\\', version)
+            productdir = Reg.get_value(key, "installdir")
+        except KeyError:
+            productdir = None
+
+    if productdir:
+        vcvarsall = os.path.join(productdir, "vcvarsall.bat")
+        if os.path.isfile(vcvarsall):
+            return vcvarsall
+    return None
+
+def _get_msvc_env(vsver, x64flag):
+    vcvars = None
+    try:
+        toolsdir = os.environ['VS%sCOMNTOOLS' % vsver]
+    except KeyError:
+        # try to import from the registry, as done in setuptools
+        # XXX works for 90 but is it generalizable?
+        toolsdir = ''
+        vcvars = _find_vcvarsall(vsver/10)
+
+    if not vcvars:
+        if x64flag:
+            vsinstalldir = os.path.abspath(os.path.join(toolsdir, '..', '..'))
[…]

### 2018-03-22 test_env/run_test.py

@@ -5,6 +5,7 @@ from httperfpy import Httperf, HttperfParser
 import matplotlib.pyplot as plt
 import json
 from threading import Thread
+import sys
 
 measurements = defaultdict(lambda: {})
 num_requests = 0
@@ -63,7 +64,7 @@ def test_load(requests=30):
     a.join()
     b.join()
 
-def plot_results():
+def plot_results(measurements):
     fig, (shared, private) = plt.subplots(2, 1)
 
     Xs = []
@@ -79,24 +80,29 @@ def plot_results():
     shared.set_xlabel("num_requests")
     shared.set_ylabel("memory (kb)")
     shared.set_title("Shared memory")
+    shared.set_xlim(left=0)
 
     private.scatter(Xs, private_Ys)
     private.set_xlabel("num_requests")
     private.set_ylabel("memory (kb)")
     private.set_title("Nonshared memory")
+    private.set_xlim(left=0)
+    private.set_ylim(bottom=0)
 
     fig.tight_layout()
     plt.show()
 
-for _ in range(30):
+if __name__ == '__main__':
+    print sys.argv[1]
+    for _ in range(60):
+        measure_all_pids()
+        test_load()
+        print _
+
     measure_all_pids()
-    test_load()
-    print _
+    pprint(dict(measurements))
 
-measure_all_pids()
-pprint(dict(measurements))
+    with open(sys.argv[1], 'w') as outfile:
+        json.dump(measurements, outfile)
 
-with open('results.json', 'w') as outfile:
-    json.dump(measurements, outfile)
-
-plot_results()
+    plot_results(measurements)

### 2018-03-23 paper/main.tex
Wrote first draft of entire paper.  Added charts

@@
 \usepackage[margin=1in]{geometry} 
+\usepackage{listings}
 
-%\usepackage{titlesec}
+\usepackage{color}
 
-%\titleformat*{\section}{\normalsize\bfseries}
-%\titleformat*{\subsection}{\normalsize\bfseries}
-%\titleformat*{\subsubsection}{\normalsize\bfseries}
-%\titleformat*{\paragraph}{\normalsize\bfseries}
-%\titleformat*{\subparagraph}{\normalsize\bfseries}
+\definecolor{codegreen}{rgb}{0,0.6,0}
+\definecolor{codegray}{rgb}{0.5,0.5,0.5}
+\definecolor{codepurple}{rgb}{0.58,0,0.82}
+\definecolor{backcolour}{rgb}{0.95,0.95,0.92}
+
+\usepackage{fontspec}
+\usepackage{pgfplots}
+\newlength\figurewidth
+\newlength\figureheight
 
@@
     \begin{abstract}
-        Most research into garbage collection is focused on analyzing a single process in isolation.
-This paper focuses on the garbage collection of many processes with shared memory.
-We examine a deployment of a Django webserver (written in python) running in multi process mode and profile its memory usage with various garbage collection algorithms.
-We then modify pypy's generational garbage collector to improve its use of copy-on-write memory.
-Specifically, we make references between objects into indirect references and store those in densely packed reference tables.
-That way when an object is relocated from a younger into an older generation, only the reference table needs to be updated to reflect the object's new location.
-This prevents the garbage collector from unnecessarily writing to shared memory.
+        Most research into Garbage Collection is focused on analyzing a single process in isolation.
+This paper focuses on the Garbage Collection of many processes with shared memory.
+We examine a deployment of a Django webserver (written in Python) running with multiple forked workers and profile its memory usage with various Garbage Collection strategies.
+We use PyPy, a Python runtime environment with a modular codebase, and modify the Garbage Collector to improve its use of copy-on-write memory.
+Specifically, we take Garbage Collector state which was previously stored inside the objects and instead store that state in separate memory pages.
+That way when a collection event occurs, those pages which were shared between Python processes can remain untouched during collection.
+Our analysis found that the modification incurs some runtime overhead which initially increases page load latency but that it also decreases memory usage in worker processes by up to 20\% and preserves shared memory between workers.
     \end{abstract}
@@
 \section{Introduction}
-Advances in interpretors for high level languages have enabled the use of high level languages in performance critical applications.
+Advances in interpretors for high level languages have enabled their use in performance critical applications.
 Some of the most visited websites on the Internet as of 2018 are written in high level languages such as Python, Ruby, and PHP.
-A common problem with these deployments is memory usage: garbage collected languages tend to use more memory than their manually-memory-managed counterparts.
+A common problem with these deployments is memory usage: Garbage Collected languages tend to use more memory than their manually-memory-managed counterparts.
 In order for these popular websites to continue to scale up, the tools on which they are developed need to get more efficient.
-Webservers need to use less memory.
+Many of these systems are memory bound and so in order to make them more efficient we must make them use less memory.
 
-Most research into garbage collection focuses on the process in isolation.
-This paper, however, examines garbage collection in a group of processes which share memory in a Copy-On-Write fashion.
+Most research into Garbage Collection focuses on the process in isolation.
+This paper, however, examines Garbage Collection in a group of processes which share memory in a Copy-On-Write fashion.
 In a webserver environment where worker processes are often forked repeatedly during operation, this situation is very common.
 Workers forked from the master process initially share some memory with the master and so the memory usage of the system as a whole is directly related to how much memory can be shared between processes.
-This is the type of environment that many of today's largest websites like Instagram\cite{dismissing_garbage} operate with and so we believe that improvements made in this area can be used to improve performance in the field as a whole.
+This is the type of environment that many of today's largest websites like Instagram operate with\cite{dismissing_garbage}.
 
-To accomplish this goal, this paper examines an optimization in PyPy's Generational Garbage Collector which is designed to reduce the number of shared pages that get dirtied during the copying phase of a minor collection.
-Section \ref{sec:background} explains various topics relating to the optimization and builds up to a demonstration intended to create some intuition into how this optimziation is designed to work.
+Oftentimes in environments like this, the Garbage Collector will disturb shared pages during collection.
+While Garbage Collection has the intention of reducing the number of pages used by a process, by disturbing shared pages in forked workers, the Garbage Collector ends up increasing memory usage in the group as a whole \cite{dismissing_garbage}.
+
+This paper examines an optimization in PyPy's Generational Garbage Collector\cite{pypy-doc} which is designed to reduce the number of shared pages that get dirtied during a major collection.
+To accomplish this, we take Garbage Collector state which was previous stored within each object and we move that state into a dense index stored in its own page in memory.
+This prevents the Garbage Collector from needing to write to shared memory during collection.
+
+Section \ref{sec:background} explains various topics relating to the optimization and builds up to a demonstration intended to create some intuition into how this optimization is designed to work.
 Section \ref{sec:design} explains the nitty gritty details of the optimization and builds intuition as to which data structures need to be changed.
-Section \ref{sec:implementation} summarizes the source changes made to PyPy's garbage collector in order to implement this.
+Section \ref{sec:implementation} summarizes the source changes made to PyPy's Garbage Collector in order to implement this.
 Section \ref{sec:method} talks about how these changes were tested and section \ref{sec:analysis} examines the results of these tests and explains them.
@@
 
-This section provides some background information that may help to contextualize the methods used in this paper.
-Section \ref{sec:CoW} is a quick review of the Linux memory model as well as an explanation of Copy-on-Write semantics.
-Sections \ref{sec:RCGC} and \ref{sec:MSGC} are a quick review of the Reference-Counting and Mark-and-Sweep algorithms used together in CPython's Garbage Collector.
-Section \ref{sec:CGGC} is an overview of the Compacting Generational Garbage Collection algorithm used in PyPy and racket.
-Finally, section \ref{sec:CoWGC} is an overview of how these algorithms can work against each other to increase memory used by a collection of related processes.
-Those familiar with the algorithms covered in sections \ref{sec:CoW}, \ref{sec:MSGC}, and \ref{sec:CGGC} may wish to skip over those sections.
-Section \ref{sec:CoWGC} is more than simple review and thus should not be skipped.
+Before presenting the modifications made to PyPy's Garbage Collector, we first motivate the need for Copy-on-Write-Sensitive Garbage Collection by explaining how shared memory gets wasted.
+First we review the Linux memory model and Copy-on-Write semantics.
+Next we review Reference Counting Garbage Collection as well as Mark-and-Sweep Garbage Collection and Compacting Generational Garbage Collection.
+Finally we put this information together to demonstrate how these algorithms when used in conjunction can end up unnecessarily wasting memory in a shared-memory environment.
 
@@
 A single address space can be shared among $N$ processes.
-Since each page has an unsigned int dedicated to reference count, quite a few pages can refer to a single page and copying will happen as needed \cite{kernelsource,VMM}.
+Since each page has an entire word dedicated to reference count, quite a few pages can refer to a single page and copying will happen as needed \cite{kernelsource,VMM}.
 
@@
 When references to this object are updated or when objects which refer to this object are collected, the reference count attribute is decremented.
-When the reference count of an attribute reaches zero, it is garbage collected and all the objects to which it refers are visited accordingly \cite{GC-continuum}.
+When the reference count of an attribute reaches zero, it is Garbage Collected  and the object is deleted in place.
+When an object is collected, all the objects that it referred to are updated visited accordingly \cite{GC-continuum}.
 
-This garbage collection algorithm can be visualized as a directed graph where every node is an object in memory and where every edge is a reference from one object to another.
-The reference count for each object is the the in-degree of its associated node.
-When there are no edges into a node, the algorithm knows that that node (and all its out-edges) can be collected.
+This Garbage Collection algorithm can be visualized as a directed graph where every node is an object in memory and where every edge is a reference from one object to another.
+The reference count for each object is the in-degree of its associated node.
+When there are no edges into a node, the algorithm knows that this node (and all its out-edges) can be collected.
 This scheme has advantages in that many objects can be collected as soon as they become unreachable.
 However, if a group of unreachable objects form a reference cycle, their reference counts will never reach zero and so they will never be collected by a strictly reference-counted system.
+Additionally, because objects are collected in place, this collection strategy is subject to fragmentation over time \cite{GC-continuum}.
 
-This algorithm has been used extensively in CPython since python 2 and accounts for the majority of object collection in most running systems \cite{cpython3-doc}.
+This algorithm has been used extensively in CPython since Python2 and accounts for the majority of object collection in most running systems \cite{cpython3-doc}.
 Whenever a reference is created, updated, or deleted, CPython visits the objects being referenced and updates their reference counts appropriately.
@@
 
-Mark-and-Sweep Garbage Collection can be modeled as a traversal of the object graph starting with a set of "accessible roots".
+Mark-and-Sweep Garbage Collection can be modeled as a traversal of the object graph starting with a set of ``accessible roots''.
 These accessible roots are objects that are being directly referenced by the program (e.g.
 globals, variables on the stack, intermediate values of a computation, etc.).
-Starting with these nodes, the algorithm recursively visits adjacent nodes marking them as "accessible" when they are visited.
-At the end of this first sweep, the algorithm has divided all objects into two sets: those which are accessible to the program (nodes tags as "accessible") and those which are not accessible to the program (nodes not tagged as "accessible").
+Starting with these nodes, the algorithm recursively visits adjacent nodes marking them as ``accessible'' when they are visited.
+At the end of this first sweep, the algorithm has divided all objects into two sets: those which are accessible to the program (nodes tagged as ``accessible'') and those which are not accessible to the program (nodes not tagged as ``accessible'').
 To do the collection, the algorithm makes a pass over every object in the system.
-Those not tagged as "accessible" are deleted in place \cite{GC-continuum}.
+Those not tagged as ``accessible'' are deleted in place \cite{GC-continuum}.
 
@@
 This increases the number of physical pages in use and increases the cache miss rate.
+Additionally, because a complete Mark-and-Sweep collection needs to traverse the entire object-graph, there is a significant pause while collection occurs \cite{GC-continuum}.
 
-This algorithm has been used in conjunction with Reference Counting in CPython since python 2.
-While the reference counter is always running in the background during CPython execution, CPython will initiate a MSGC (usually called a "major collection event") collection once memory usage exceeds a certain threshold.
+This algorithm has been used in conjunction with Reference Counting in CPython since Python2.
+While the reference counter is always running in the background during CPython execution, CPython will initiate a MSGC (usually called a ``major collection event'') collection once memory usage exceeds a certain threshold.
 In combination, these algorithms eliminate most garbage in a timely manor while also correctly collecting cycles that form on occasion \cite{cpython3-doc}.
@@
 
-In Compacting Generational Garbage Collection, objects are divided into multiple "generations" which are each located on separate physical pages of memory.
+In Compacting Generational Garbage Collection, objects are divided into multiple ``generations'' which are each located on separate physical pages of memory.
 In many cases there are 3 or 4 successively larger generations.
-New objects are inserted into generation 0 (often called the "nursery").
+New objects are inserted into generation 0 (often called the ``nursery'').
 Once generation 0 reaches a certain size, the objects in generation 0 are traversed starting with the accessible roots.
@@
 
-In most implementations there is a final generation (often called the "mature" generation) for objects that have survived all the prior collections.
-There are several ways to handle this final generation - the most common being to let the garbage build up.
+In order to correctly account for inter-generational references, each generation must keep track of references from other generations into itself.
+When an object in generation $M$ is updated to reference an object in generation $N$, that object is added to generation $N$'s cross-reference list.
+When generation $N$ is collected, the Garbage Collector adds references in the cross-reference list to the root accessible set and those references are used when tracing which objects are in use.
+When an object in generation $M$ is copied to generation $M + 1$, all the updates to that object must be updated to refer to this new location in memory \cite{GC-continuum}.
+
+In most implementations there is a final generation (often called the ``mature'' generation) for objects that have survived all the prior collections.
+There are several ways to handle this final generation - the most common being to simply never collect it.
 This implementation is not technically correct but it performs well in practice.
-In PyPy, the mature generation is collected using the MSGC.
-Memory fragmentation which may occur in PyPy's mature generation are possible filled when objects are copied from younger generations into mature \cite{pypy-doc}.
 
-\begin{quotation}
-    "Many generational GC's are not comprehensive--they don't successfully remove all the garbage (long-lived garbage, in particular, may never get collected).
-But it's the fresh garbage that smells the worst.
-:)"
-
-    -- C2 Wiki\cite{C2}
-\end{quotation}
-
-\subsubsection{Inter-Generational References: Always Use Protection!}
-
-The CGGC algorithm as described above is incomplete.
-It's possible for objects in one generation to refer to objects in another generation and doing a naive "accessible-roots" seeded traversal will not see these references which may cause some objects to get collected prematurely.
-To prevent this we introduce a protective write barrier.
-
-Every generation keeps a list of cross-generational references into that generation.
-When an object in generation $M$ is updated to reference an object in generation $N$, that object is added to generation $N$'s cross-reference list.
-When generation $N$ is collected, the garbage collector adds references in the cross-reference list to the root accessible set and those references are used when tracing which objects are in use.
-When an object in generation $M$ is copied to generation $M + 1$, all the updates to that object must be updated to refer to this new location in memory \cite{GC-continuum}.
+The PyPy Python environment uses CGGC with two generations: a ``nursery generation'' for young objects and a ``mature generation'' for mature objets.
+On occasion, the mature generation is collected using the MSGC.
+Memory fragmentation in the mature generation can be filled when objects are copied from younger generations into mature \cite{pypy-doc}.
 
@@
 Each of these algorithms alone serve to minimize the memory used by an individual process as it runs.
-CoW allows related processes to share physical memory, and all three garbage collection algorithms reduce the number of physical pages needed by each process.
-However, when used in conjunction we often see that the memory performance of a collection of related processes suffers.
+CoW allows related processes to share physical memory, and all three Garbage Collection algorithms reduce the number of physical pages needed by each process.
+However, when used in conjunction we often see that the memory performance of a collection of related processes suffers \cite{dismissing_garbage}.
 
@@
 When this happens, the reference count field in the referenced object must be updated.
-This triggers a copy-on-write because even though none of of the objects stored on the page changed, the meta-data that the garbage collector stores on the page has changed.
+This triggers a copy-on-write because even though none of of the objects stored on the page changed, the meta-data that the Garbage Collector stores on the page has changed.
 In this way, reference counting breaks shared pages which can increase memory used by the system.
 
-When a CPython major collection event occurs, the MSGC algorithm will start a full heap traversal that marks objects as "accessible" or "not-accessible".
-Because the garbage collector stores this information adjacent to the objects in memory, the shared pages containing these objects get written to which triggers a copy-on-write and the page sharing breaks.
-In this way, a single major garbage collection event can greatly increase memory used by the system.
+When a CPython major collection event occurs, the MSGC algorithm will start a full heap traversal that marks objects as ``accessible'' or ``not-accessible''.
+Because the Garbage Collector stores this information adjacent to the objects in memory, the shared pages containing these objects get written to which triggers a copy-on-write and the page sharing breaks.
+In this way, a single major Garbage Collection event can greatly increase memory used by the system.
 
-This problem is somewhat alleviated when using a generational garbage collector such as is included in PyPy because we can copy objects out of a shared page without writing to it, however when we do the copy, those inter-generational references must be updated.
-Doing a minor collection of generation 1 may force writes to generations 2, 3, 4, etc, which breaks the sharing of those pages.
+This problem is somewhat alleviated when using CGGC such as is included in PyPy because we can copy objects out of a shared page without writing to it, however PyPy's GC is also subject to the dirtying of pages that happens when the mature generation is Mark-and-Swept.
 
@@
 
-The default garbage collector in PyPy is called "minimark" and it is a classic Generational Garbage Collector with 2 generations \cite{pypy-doc}.
-New objects are created in the nursery and are moved to gen1 on minor collection events.
+In this paper we modify the Garbage Collector in the PyPy Python runtime environment.
+We choose PyPy because it is written in a dialect of Python which makes the interpreter easy to modify for experimental purposes.
+Also, because PyPy's Garbage Collector is a relatively simple CGGC, the modification and analysis is relatively straightforward.
+
+To prevent the Garbage Collector from touching objects in place, we add a hashtable to the Garbage Collector that contains the set of all the elements that have been visited.
+When a Mark-and-Sweep occurs in the mature generation, rather than writing to the objects in place, the Garbage Collector inserts the object's id into the hashtable.
+When the Mark-and-Sweep finishes, that hashtable is cleared.
+
+\section{Implementation}\label{sec:implementation}
+
+PyPy has several Garbage Collectors but the one we chose to patch for the purposes of this paper is called ``minimark''.
+Minimark is a CGGC with two generations \cite{pypy-doc}.
+New objects are created in the ``nursery'' and are moved to the ``mature'' generation on minor collection events.
+When the mature generation exceeds a certain size, a major collection starts and the mature generation is Mark-and-Swept.
+During this MSGC, each object in the mature generation is visited and a bit is set in-place which tags that object as ``safe'' during this collection.
+
+To implement this change, we added a member variable to the MiniMarkGC class of type $rpython.memory.support.AddressDict$.
+PyPy's memory library does not provide an AddressSet type so this was chosen as an approximation.
+Objects which are in the mapping and which map to themselves are considered visited and objects which either are not in the mapping or which map to NULL are considered unvisited.
+
+The following code snippet was added to MiniMarkGC and demonstrates the new interface for determining an object's visited state.
+
+\begin{lstlisting}[language=python]
+def obj_visited(self, obj):
+  return (
+    self.visited_objs.contains(obj) and
+    self.visited_objs.get(obj) == obj)
+
+def mark_obj_visited(self, obj):
+  self.visited_objs.setitem(obj, obj)
+
+def unmark_obj_visited(self, obj):
+  self.visited_objs.setitem(obj,
+        llmemory.NULL)
+\end{lstlisting}
+
+\section{Methodologies}\label{sec:method}
+
+This paper focuses on a web-server workload and so to validate our patch, we chose to create the closest approximation of a production webserver possible.
+Our model was Instagram's web-tier servers which are briefly described in their external blog.
+Instagram is a massive Django instance running atop CPython2.7 using Ngnx as a front-end server and using uWSGI to manage Django worker processes \cite{dismissing_garbage}.
+uWSGI is running in pre-fork mode meaning that a worker is started and then forked several times to create more workers.
+When a worker dies, an existing worker is forked to replace it.
+
+With this test environment we create 2 experimental groups.  The groups are as follows:
 
 \begin{enumerate}
-    \item Currently when mature objects refer to old objects, that reference is stored in a special "write barrier" so their references are counted
-    \item When young object moves from young to mature generation, the objects which refer to it get updated in-place
-    \item We are going to put all references between objects in a reference table... every lookup will be a double lookup but every update will be a single edit to the table
+    \item Minimark GC without patch
+    \item Minimark GC with patch
 \end{enumerate}
 
-\section{Implementation}\label{sec:implementation}
-\begin{enumerate}
-    \item Give a summary of which structures and functions actually changed
-    \item Where did I store this information?  How is it accessed in code?  
-    \item maybe some pseudocode snippets
-\end{enumerate}
+\subsection{Test Environment}
 
-\section{Methodologies}\label{sec:method}
-\begin{enumerate}
-    \item Ngnx + uWSGI + Django + pypy in multi-process mode
-    \item Use thebluealliance.com as a test load because it caches a lot of data and does some non-trivial computation
-    \item Generate a bunch of fake traffic using httperf
-    \item Monitor running code
-        \begin{enumerate}
-            \item How long did it take to handle 10,000 requests
-            \item What was the average time-to-last-byte?
-            \item Total memory of all processes
-            \item Average memory of each process
-            \item shared memory per process
-            \item num page faults
-            \item cache miss rate?
-            \item pages served per second (did it go up or down?)
-            \item How many times did major and minor garbage collection happen?
-            \item How many times did this inter-generation optimization actually occur?
-        \end{enumerate}
-\end{enumerate}
+In order to replicate this environment, we created a docker image that had Ngnx, uWSGI, Django, and PyPy installed on it.
+We ran uWSGI in pre-fork mode with 4 worker processes.
+The docker image was capped at 8GB of memory and 3 cpu cores.
+All benchmarks were ron on a ThinkPad T460 with 12GB ram and an i5 2.4GHz quad-core processor.
+Because the docker image was capped at well below the host's resource capacity, we believe benchmarks are reasonably isolated from resource contention with host processes.
+
+Because Instagram is closed-source, proprietary software, we were unable to procure its source code for use in testing.
+Instead we created a Django application which was created specifically to emulate a production web-server load.
+Specifically, we created a Django application which allocates a large chunk of memory at startup and which allocates an additional chunk of memory for each request.
+Below are the snippets of code used to create test-garbage:
+
+\begin{lstlisting}[language=python]
+class LargeObject():
+  # Use a bunch of memory
+  def __init__(self):
+    self.lists = []
+    self.strs = []
+    for i in range(1600):
+      self.lists.append(['text'] * i)
+      for j in range(40):
+        self.strs.append('str' * 8)
+
+def on_startup():
+  global static_memory
+  static_memory = [
+    LargeObject()
+    for x
+    in range(5)]
+
+def on_page_load():
+  global glob_a, glob_b, glob_c
+  glob_a, glob_b, glob_c = (
+    LargeObject(), glob_a, glob_b)
+\end{lstlisting}
+
+We believe that this mix of memory allocation somewhat approximates what might be seen in a production web-server.
+
+\subsection{Stress-Testing the Deployment}
+
+To generate a test load we used httperf which makes a series of http requests to a given http endpoint.
+We chose to track two specific metrics for each experimental group: The time it took to handle $N$ requests and the memory used after handling $N$ requests.
+
+To determine the time to handle $N$ requests, we use the httperf tool to make $N$ requests and record the wall-time it took for all requests to get a response.
+We run this experiment with $N$ values of 5k, 10k, 15k, and 20k.
+
+To track memory usage as the number of requests increases, we make a burst of 100 requests and then check the memory usage of each uWSGI worker process.
+We repeat this process until we've made 50 bursts of requests or 5000 requests in total.
+At each step we check memory usage by scraping that process's $/proc/PID/smaps$ file.
+The private memory used by that process is the sum of all lines starting with ``Private\_clean'' and ``Private\_dirty'' and the public memory used by that process is the sum of all lines starting with ``Public\_clean'' and ``Public\_dirty''.
 
 \section{Analysis}\label{sec:analysis}
-\begin{enumerate}
-    \item Did we actually save any memory?  How much on average?  Were there any patterns?   
-    \item How often did old objects actually refer to young objects?  
-    \item How often did minor collection happen?  Major collection?
-    \item How much CPU overhead was there in following that extra pointer every time?  Can I even measure this?
-    \item How many pages did we serve per second?  Did it go up or down?
-\end{enumerate}
+
+Experimental results show that this approach was successful in preserving shared memory in each worker process.
+During our verification step we measured two types of performance:  memory usage and latency.
+In the first section here we see that the patched version of the minimark GC uses less memory overall than the Standard version by preserving shared memory.
+In the second section here we see that the Standard version serves more requests per second up until a certain point after which it underperforms because of paging.
+This implies that while there is significant runtime overhead in this optimization, the costs can be outweighed by the savings in memory.
+
+\subsection{Memory Usage}
+
+Figures \ref{fig:patched_mem} and \ref{fig:standard_mem} clearly show that the patched version of the Garbage Collector preserves more shared memory than the unpatched version.
+In the upper portion of each figure we see the shared memory for each case.
+In the Standard GC group, we see that the amount of shared memory trails off as the number of requests grows because when each of the worker processes runs major Garbage Collection they taint many of the pages.
+On the other hand, the shared memory chart for the Patched GC is almost entirely flat meaning that those shared pages are untouched and preserved.
+In the lower chart we can see that after 5000 requests, the Patched worker processes take up somewhere in the realm of 800Mb of space per worker while the Standard worker processes take up 1000 to 1200Mb of space per worker.
+This accounts for a reduction of around 20\%.
+
+\begin{figure}\label{fig:patched_mem}
+    \begin{center}\textbf{Memory Usage of Patched GC}\end{center}
+    \resizebox{0.5\textwidth}{!}{\input{patched_mem.pgf}}
+    \caption{Memory usage of patched Garbage Collector running 4 workers after $X$ requests.  The }
+\end{figure}
+
+\begin{figure}\label{fig:standard_mem}
+    \begin{center}\textbf{Memory Usage of Standard GC}\end{center}
+    \resizebox{0.5\textwidth}{!}{\input{standard_mem.pgf}}
+    \caption{Memory usage of standard Garbage Collector running 4 workers after $X$ requests}
+\end{figure}
+
+\subsection{Latency}
+
+The table below shows the time per request for our deployment with and without the optimization.
+Figure \ref{fig:latency} shows the same information in a graphical format.
+We can see that for a small amount of traffic, the standard minimark GC responses to more requests per second than our patched version.
+However, once the number of requests per second exceeded around 13k, the total memory of the system exceeded the 8GB allotted and the deployment started paging aggressively.
+On the other hand the patched GC continues to settles down for larger numbers of requests.
+The upward trend in response rate is a result of PyPy's tracing JIT'er - over the course of a program's lifetime PyPy will compile commonly executed codepaths.
+When the application is freshly started, PyPy interprets the source code and spends time compiling codepaths.
+After the application has been running for a while, PyPy spends less time interpreting and compiling and more time executing native code \cite{pypy-doc}
+
+\begin{center}
+\begin{tabular}{| l || l | l |}
+    \hline
+     & \textbf{Standard} & \textbf{Patched} \\
+    \hline
+    5k Requests & 190.8 conn/sec & 67.9 conn/sec \\
+    10k Requests & 269.2 conn/sec & 148.7 conn/sec \\
+    15k Requests & 97.7 conn/sec & 188.8 conn/sec \\
+    20k Requests & 103.2 conn/sec & 190.9 conn/sec \\
+    \hline
+\end{tabular}
+\end{center}
+
+From this data we can see that there is time overhead in our optimization.
+If we had allocated more than 8GB of memory to the test, the Standard GC process would likely not have paged and would have responded to more requests per second than the patched version.
+However the patched version used less memory overall and thus did not spend any time paging.
+
+\begin{figure}\label{fig:latency}
+    \resizebox{0.5\textwidth}{!}{\input{chart_latency.pgf}}
+    \caption{Request Rate for different Request Quantities.  Note Standard GC started paging after 10k requests}
+\end{figure}
 
 \section{Future Work}\label{sec:futurework}
-\begin{enumerate}
-    \item What other changes could I make to improve CoW performance?  Lemme think.
-\end{enumerate}
+
+While the optimization explored in this paper account for noticeably memory savings, there is much more that can be done in the future.
+
+In a CGGC like PyPy, when objects are copied from Eden into the mature generation, all the mature objects which pointed to the new object in Eden must be updated to refer to that objects new post-copy location.
+This results in writes to mature objects that can be avoided.
+If instead of updating the pointers to young objects in place in the old objects, there were a reference table so that all references from old to young objects were indirect, then when young objects were copied only the reference table would need to be updated.
+This lets the mature objects remain untouched even when the object which they are referencing moves.
+
+Additionally, the benchmark used in this paper was concocted specifically for testing purposes.
+The results from using this benchmark may or may not accurately reflect the reality of real-world Django deployments.
+Future work would likely take the findings in this paper and apply them to actual existing production websites.
+Finding an open source website that runs on Django that is relatively easy to deploy would be paramount in bringing this research to real-world relevance.
+
+Lastly it would be interesting to see how the preservation of shared memory effects the cache miss rate of the worker processes.
+In theory, if the worker processes are sharing physical memory then a single line-item in the cache can be shared amongst worker processes.
+Whether this happens in would be worth looking into.
 
@@
 
-This paper is very much inspired by recent updates in CPython's and Ruby MRI's garbage collectors which were updated in 2017 and 2014 respectively.
+This paper is very much inspired by recent updates in CPython's and Ruby MRI's Garbage Collectors which were updated in 2017 and 2014 respectively.
 
 \subsection{Copy-On-Write optimization in Ruby MRI's MSGC}
-Ruby 2.0 contains an update to the GC that improved CoW performance.
-Ruby's GC is a classic CGGC with a young generation and an old generation.
-In the old algorithm, every object contained a "visited" bit that would note whether the object had visited by the tracer.
-In the new GC algorithm, these "visited" bits are stored outside the objects being traced.
+
+Ruby 2.0 contains a similar update to the GC that improved CoW performance.
+Ruby's GC is a classic CGGC with a young generation and an old generation that is occasionally Mark-and-Swept.
+In the old algorithm, every object contained a ``visited'' bit that would note whether the object had visited by the tracer.
+In the new GC algorithm, these ``visited'' bits are stored outside the objects being traced.
 In this way, the objects themselves don't need to be updated during a GC event.
 
-The optimization described in this paper is different in that we focus on preserving old objects when objects are copied from Eden into the older generation.
+The optimization described in this paper is very similar to this however we are specifically testing this optimization in the context of Python for a webserver workload.
 
 \subsection{Copy-On-Write Friendly Garbage Collection in CPython at Instagram}
-A 2017 article from Instagram's Engineering series describes changes to CPython's garbage collector made at Instagram aimed at improving memory performance in their web-tier servers.
+
+A 2017 article from Instagram's Engineering series describes changes to CPython's Garbage Collector made at Instagram aimed at improving memory performance in their web-tier servers.
 The experimentation section of this paper is heavily inspired by the test setup used at Instagram.
 
-The optimization implemented at Instagram was to have the programmer manually mark certain objects as "GC-Frozen".
-Objects that are GC-Frozen are not touched by the garbage collector and so the data in and around them is not touched.
+The optimization implemented at Instagram was to have the programmer manually mark certain objects as ``GC-Frozen''.
+Objects that are GC-Frozen are not touched by the Garbage Collector and so the data in and around them is not touched.
 Because Instagram's specific use case involves a lot of static objects that are shared between processes, this worked effectively.
+
+\section{Conclusion}
+
+The results in this paper clearly show that by moving Garbage Collector state out of the objects and into isolated Garbage Collector memory, we can prevent unnecessary memory writes and we can preserve shared memory between forked processes.
+While the changes described in this paper have some runtime overhead in terms of mapping objects in memory to their associated Garbage Collector state, we found that for our deployment the savings in memory far outweighed the runtime overhead of the change.
 

### 2018-03-23 paper/main.tex
Roll in feedback on paper

@@
 
-\pretolerance=10000
+\pretolerance=100
 
@@
 \begin{document}
+\begin{sloppypar}
 
@@
 We use PyPy, a Python runtime environment with a modular codebase, and modify the Garbage Collector to improve its use of copy-on-write memory.
-Specifically, we take Garbage Collector state which was previously stored inside the objects and instead store that state in separate memory pages.
+Specifically, we take the Garbage Collector state which was previously stored inside the objects and instead store that state in separate memory pages.
 That way when a collection event occurs, those pages which were shared between Python processes can remain untouched during collection.
-Our analysis found that the modification incurs some runtime overhead which initially increases page load latency but that it also decreases memory usage in worker processes by up to 20\% and preserves shared memory between workers.
+Our analysis found that the modification decreases private memory usage in worker processes by up to 20\% and preserves shared memory between workers.
     \end{abstract}
@@
 \section{Introduction}
-Advances in interpretors for high level languages have enabled their use in performance critical applications.
+Advances in interpreters for high level languages have enabled their use in performance critical applications.
 Some of the most visited websites on the Internet as of 2018 are written in high level languages such as Python, Ruby, and PHP.
@@
 
-Oftentimes in environments like this, the Garbage Collector will disturb shared pages during collection.
+Oftentimes in environments like this, the Garbage Collector will disturb pages during collection by writing to them.
+When doing a Mark-and-Sweep operation, the Garbage Collector traverses the entire object-space and marks the objects which are live.
+This marking disturbs the page regardless of whether any objects within it are actually garbage.
 While Garbage Collection has the intention of reducing the number of pages used by a process, by disturbing shared pages in forked workers, the Garbage Collector ends up increasing memory usage in the group as a whole \cite{dismissing_garbage}.
 
-This paper examines an optimization in PyPy's Generational Garbage Collector\cite{pypy-doc} which is designed to reduce the number of shared pages that get dirtied during a major collection.
-To accomplish this, we take Garbage Collector state which was previous stored within each object and we move that state into a dense index stored in its own page in memory.
+This paper examines an optimization in PyPy's Generational Garbage Collector\cite{pypy-doc} which is designed to reduce the number of shared pages that get disturbed during a major collection.
+To accomplish this, we take the Garbage Collector state which was previously stored within each object and we move that state into a dense index stored in its own page in memory.
 This prevents the Garbage Collector from needing to write to shared memory during collection.
@@
 Section \ref{sec:background} explains various topics relating to the optimization and builds up to a demonstration intended to create some intuition into how this optimization is designed to work.
-Section \ref{sec:design} explains the nitty gritty details of the optimization and builds intuition as to which data structures need to be changed.
+Section \ref{sec:design} explains the details of the optimization and builds intuition as to which data structures need to be changed.
 Section \ref{sec:implementation} summarizes the source changes made to PyPy's Garbage Collector in order to implement this.
@@
 When a process in Linux forks, a new process is spawned with an identical address space.
-In order to save memory, the Linux kernel does not copy the entire address space, but rather it copies the page table of the parent process and marks all physical pages as shared and read-only \cite{VMM}.
+In order to save memory, the Linux kernel does not copy all the physical pages in the parent process, but rather it copies the page table of the parent process and marks all physical pages as shared and read-only \cite{VMM}.
 If the physical page was already marked shared, the reference count of the page is incremented.
@@
 If the physical page is still being shared, the kernel allocates a new physical page, copies the contents of the shared page into the new physical page, decrements the reference count in the shared page, and inserts the new physical page into the faulted process's page table \cite{VMM}.
-If the physical page is no longer being shared (for example if the child process already wrote to the physical page in question and made their own copy) then the page is converted from a read-only shared page to a read-write private page \cite{VMM}.
+If the original physical page is no longer being shared (for example if all the other processes already made their own copies) then the page is converted from a read-only shared page to a read-write private page \cite{VMM}.
 In this way, memory is allocated and the process's page-table is updated in an extremely lazy manor which has proven to be very efficient in practice.
@@
 When the reference count of an attribute reaches zero, it is Garbage Collected  and the object is deleted in place.
-When an object is collected, all the objects that it referred to are updated visited accordingly \cite{GC-continuum}.
+When an object is collected, all the objects that it referred to are updated accordingly \cite{GC-continuum}.
 
@@
 
-A common problem with this algorithm is that objects are deleted in-place and so the physical memory where objects reside tends to become fragmented over time.
+Much like with RCGC, a common problem with this algorithm is that objects are deleted in-place and so the physical memory where objects reside tends to become fragmented over time.
 This increases the number of physical pages in use and increases the cache miss rate.
-Additionally, because a complete Mark-and-Sweep collection needs to traverse the entire object-graph, there is a significant pause while collection occurs \cite{GC-continuum}.
+Unlike the RCGC, because a complete Mark-and-Sweep collection needs to traverse the entire object-graph, there is a significant pause while collection occurs \cite{GC-continuum}.
 
@@
 In Compacting Generational Garbage Collection, objects are divided into multiple ``generations'' which are each located on separate physical pages of memory.
-In many cases there are 3 or 4 successively larger generations.
 New objects are inserted into generation 0 (often called the ``nursery'').
@@
 
-When generation 1 gets larger than some threshold, it is collected in the same way and accessible objects are copied to generation 2.
+In many cases there are 3 or 4 successively larger generations.
+The same rules for moving objects from generation 0 to generation 1 apply.
 Indeed when generation $N$ gets large enough, its contents are traversed, accessible objects are copied to generation $N+1$, and generation $N$ is cleared.
@@
 There are several ways to handle this final generation - the most common being to simply never collect it.
-This implementation is not technically correct but it performs well in practice.
+This implementation is not technically correct because mature objects which become garbage will never be collected but it performs well in practice.
 
@@
 
-\subsection{Tragedy of the commons: Dirtying Shared Pages in CGGC}\label{sec:CoWGC}
+\subsection{Tragedy of the commons: Dirtying Shared Pages During GC}\label{sec:CoWGC}
 
@@
 
-As a thought experiment, consider a forking webserver written in python running atop a standard CPython interpretor.
+As a thought experiment, consider a forking webserver written in Python running atop a standard CPython interpreter.
 When a request comes in, the process forks to spawn a worker thread which begins processing the request.
@@
 This triggers a copy-on-write because even though none of of the objects stored on the page changed, the meta-data that the Garbage Collector stores on the page has changed.
-In this way, reference counting breaks shared pages which can increase memory used by the system.
+In this way, RCGC disturbs shared pages which can increase memory used by the system.
 
@@
 Because the Garbage Collector stores this information adjacent to the objects in memory, the shared pages containing these objects get written to which triggers a copy-on-write and the page sharing breaks.
-In this way, a single major Garbage Collection event can greatly increase memory used by the system.
+In this way, MSGC disturbs shared pages which can increase memory used by the system.
 
@@
 
-While these algorithms are successful at reducing the memory usage of an individual process, when examining a collection or related processes they do poorly at reducing total combined memory usage.
-The next section is dedicated to modifying CGGC to solve this problem.
+While these algorithms are successful at reducing the memory usage of an individual process, when examining a collection of related processes they do poorly at reducing total combined memory usage.
+The next section is dedicated to modifying PyPy's CGGC to prevent the Mark-and-Sweep phase from disturbing shared pages.
 
@@
 Also, because PyPy's Garbage Collector is a relatively simple CGGC, the modification and analysis is relatively straightforward.
+PyPy's Garbage Collector is a CGGC with generations: a ``nursery generation'' for young objects and a ``mature generation'' for mature objets.
+On occasion, the mature generation is collected using the MSGC.
+Copying Generational Garbage Collection events are referred to as ``minor collections'' while Mark-and-Sweep Garbage Collection events are referred to as ``major collections'' \cite{pypy-doc}.
+The modification described here is intended only to affect major collections.
 
@@
 When the Mark-and-Sweep finishes, that hashtable is cleared.
+This hashtable will be stored in different physical pages from the objects which it describes.
+By doing this we create two categories of pages: those that contain objects and those that contain object metadata.
+The GC usually only needs to modify the pages that contain object metadata which means that the pages containing objects can remain undisturbed.
 
@@
 PyPy's memory library does not provide an AddressSet type so this was chosen as an approximation.
-Objects which are in the mapping and which map to themselves are considered visited and objects which either are not in the mapping or which map to NULL are considered unvisited.
+Objects which are in the mapping and which map to themselves are considered visited and objects which either are not in the mapping or which map to $rpython.memory.Null$ are considered unvisited.
 
@@
 \end{lstlisting}
+
+We treat $visited\_objs$ like a set.
+At the start of collection, $visited\_objs$ is initialized to be an empty set.
+During the Mark phase, objects are inserted into this set.
+During the sweep phase, the entire object-space is scanned again.
+When objects are traversed, they are either collected if they were not visited during the Mark phase or are unmarked if they were visited in the Mark phase.
+Afterwards, if running in debug mode, PyPy traverses the entire object-space a third time to ensure consistency.
+In a clasic MSGC, the third pass is unnecessary, but PyPy does it to simplify development.
 
@@
 The docker image was capped at 8GB of memory and 3 cpu cores.
-All benchmarks were ron on a ThinkPad T460 with 12GB ram and an i5 2.4GHz quad-core processor.
+All benchmarks were ron on a ThinkPad T460 with 12GB ram and an i5-6300 2.4GHz quad-core processor.
 Because the docker image was capped at well below the host's resource capacity, we believe benchmarks are reasonably isolated from resource contention with host processes.
@@
 During our verification step we measured two types of performance:  memory usage and latency.
-In the first section here we see that the patched version of the minimark GC uses less memory overall than the Standard version by preserving shared memory.
+In the first section here we see that the Patched version of the minimark GC uses less memory overall than the Standard version by preserving shared memory.
 In the second section here we see that the Standard version serves more requests per second up until a certain point after which it underperforms because of paging.
-This implies that while there is significant runtime overhead in this optimization, the costs can be outweighed by the savings in memory.
+The latency in the Patched version can be attributed to the runtime overhead of looking up objects in the visited-objects set.
+The observed latency implies that depending on the setup, the cost of the modification can be outweighed by the savings in memory.
 
@@
 
-Figures \ref{fig:patched_mem} and \ref{fig:standard_mem} clearly show that the patched version of the Garbage Collector preserves more shared memory than the unpatched version.
+Figures \ref{fig:patched_mem} and \ref{fig:standard_mem} clearly show that the Patched version of the Garbage Collector preserves more shared memory than the Standard version.
 In the upper portion of each figure we see the shared memory for each case.
@@
 
-\begin{figure}\label{fig:patched_mem}
+\begin{figure}
     \begin{center}\textbf{Memory Usage of Patched GC}\end{center}
     \resizebox{0.5\textwidth}{!}{\input{patched_mem.pgf}}
-    \caption{Memory usage of patched Garbage Collector running 4 workers after $X$ requests.  The }
+    \caption{Memory usage of patched Garbage Collector running 4 workers after $X$ requests.  Each dot represents a worker's memory consumption after $X$ requests.}\label{fig:patched_mem}
 \end{figure}
 
-\begin{figure}\label{fig:standard_mem}
+\begin{figure}
     \begin{center}\textbf{Memory Usage of Standard GC}\end{center}
     \resizebox{0.5\textwidth}{!}{\input{standard_mem.pgf}}
-    \caption{Memory usage of standard Garbage Collector running 4 workers after $X$ requests}
+    \caption{Memory usage of standard Garbage Collector running 4 workers after $X$ requests.  Each dot represents a worker's memory consumption after $X$ requests.}\label{fig:standard_mem}
 \end{figure}
@@
 Figure \ref{fig:latency} shows the same information in a graphical format.
-We can see that for a small amount of traffic, the standard minimark GC responses to more requests per second than our patched version.
+We can see that for a small amount of traffic, the standard minimark GC responds to more requests per second than our patched version.
 However, once the number of requests per second exceeded around 13k, the total memory of the system exceeded the 8GB allotted and the deployment started paging aggressively.
 On the other hand the patched GC continues to settles down for larger numbers of requests.
-The upward trend in response rate is a result of PyPy's tracing JIT'er - over the course of a program's lifetime PyPy will compile commonly executed codepaths.
+The upward trend in response rate is a result of PyPy's tracing Just-in-Time compilation - over the course of a program's lifetime PyPy will compile commonly executed codepaths.
 When the application is freshly started, PyPy interprets the source code and spends time compiling codepaths.
@@
 If we had allocated more than 8GB of memory to the test, the Standard GC process would likely not have paged and would have responded to more requests per second than the patched version.
-However the patched version used less memory overall and thus did not spend any time paging.
+However the patched version used less memory overall and thus in our setup did not spend any time paging.
 
-\begin{figure}\label{fig:latency}
+\begin{figure}
     \resizebox{0.5\textwidth}{!}{\input{chart_latency.pgf}}
-    \caption{Request Rate for different Request Quantities.  Note Standard GC started paging after 10k requests}
+    \caption{Request Rate for different Request Quantities.  Note Standard GC started paging after 10k requests}\label{fig:latency}
 \end{figure}
@@
 
-In a CGGC like PyPy, when objects are copied from Eden into the mature generation, all the mature objects which pointed to the new object in Eden must be updated to refer to that objects new post-copy location.
+In a CGGC like PyPy, when objects are copied from the nursery into the mature generation, all the mature objects which pointed to the new object in the nursery must be updated to refer to that objects new post-copy location.
 This results in writes to mature objects that can be avoided.
@@
 
-Lastly it would be interesting to see how the preservation of shared memory effects the cache miss rate of the worker processes.
+Lastly it would be interesting to see how the preservation of shared memory affects the cache miss rate of the worker processes.
 In theory, if the worker processes are sharing physical memory then a single line-item in the cache can be shared amongst worker processes.
@@
 In the new GC algorithm, these ``visited'' bits are stored outside the objects being traced.
-In this way, the objects themselves don't need to be updated during a GC event.
+In this way, the objects themselves don't need to be updated during a GC event \cite{incremental_ruby}.
 
-The optimization described in this paper is very similar to this however we are specifically testing this optimization in the context of Python for a webserver workload.
+The optimization described in this paper is very similar to this, however we are specifically testing this optimization in the context of Python for a webserver workload.
 
@@
 
-The optimization implemented at Instagram was to have the programmer manually mark certain objects as ``GC-Frozen''.
-Objects that are GC-Frozen are not touched by the Garbage Collector and so the data in and around them is not touched.
-Because Instagram's specific use case involves a lot of static objects that are shared between processes, this worked effectively.
+In their article, the Instagram engineering team explored multiple ways of reducing memory usage.
+The first modification they tried was taking the reference counts stored in every Python object and moving them to a separate page.
+The intent was that when reference counts get updated, only the reference count page needed to be changed and the page storing the objects could remain shared.
+In practice this modification did not reduce memory usage \cite{dismissing_garbage}.
+
+The modification Instagram ended up implementing was to have the programmer manually mark certain objects as ``GC-Frozen''.
+Objects that are GC-Frozen are not touched by the Garbage Collector and so the Garbage Collector state embedded within them is not touched.
+Because Instagram's specific use case involves a lot of static objects that are shared between processes, this worked effectively \cite{fixing_garbage}.
 
@@
 
-The results in this paper clearly show that by moving Garbage Collector state out of the objects and into isolated Garbage Collector memory, we can prevent unnecessary memory writes and we can preserve shared memory between forked processes.
-While the changes described in this paper have some runtime overhead in terms of mapping objects in memory to their associated Garbage Collector state, we found that for our deployment the savings in memory far outweighed the runtime overhead of the change.
+The results in this paper clearly show that by moving the Garbage Collector state out of the objects and into isolated Garbage Collector memory, we can prevent unnecessary memory writes and we can preserve shared memory between forked processes.
+While the changes described in this paper have some runtime overhead in terms of mapping objects in memory to their associated Garbage Collector state, we found that for our deployment the savings in memory far outweighed the runtime overhead.
 
@@
 
+\end{sloppypar}
 \end{document}

### 2018-03-23 paper/main.tex
Final draft of paper

@@
 \definecolor{codegreen}{rgb}{0,0.6,0}
+\definecolor{codered}{rgb}{0.6,0,0}
 \definecolor{codegray}{rgb}{0.5,0.5,0.5}
@@
 \definecolor{backcolour}{rgb}{0.95,0.95,0.92}
+
+% Python style for highlighting
+\lstdefinestyle{pystyle}{
+    language=python,
+    otherkeywords={self},
+    keywordstyle=\color{codered},
+    emphstyle=\color{codepurple},    % Custom highlighting style
+    stringstyle=\color{codegreen},
+    frame=tb,                         % Any extra options here
+    showstringspaces=false            % 
+}
+
+\lstset{style=pystyle}
 
@@
 This prevents the Garbage Collector from needing to write to shared memory during collection.
+We then provide experimental evidence showing that this optimization does in fact reduce memory usage of a group of processes.
 
@@
 
+\noindent
+\begin{minipage}{\linewidth}
 \begin{lstlisting}[language=python]
@@
 \end{lstlisting}
+\end{minipage}
 
@@
     for i in range(1600):
-      self.lists.append(['text'] * i)
+      self.lists.append(["text"] * i)
       for j in range(40):
-        self.strs.append('str' * 8)
+        self.strs.append("str" * 8)
 
@@
 At each step we check memory usage by scraping that process's $/proc/PID/smaps$ file.
-The private memory used by that process is the sum of all lines starting with ``Private\_clean'' and ``Private\_dirty'' and the public memory used by that process is the sum of all lines starting with ``Public\_clean'' and ``Public\_dirty''.
+The private memory used by that process is the sum of all lines starting with ``Private\_clean'' and ``Private\_dirty'' and the shared memory used by that process is the sum of all lines starting with ``Shared\_clean'' and ``Shared\_dirty''.
 

### 2018-03-23 test_env/chart_latency.py
Data collection

@@ -1,0 +1,25 @@
+import matplotlib.pyplot as plt
+import sys
+import json
+from matplotlib2tikz import save as tikz_save
+
+Xs = [5, 10, 15, 20]
+standard_conn = [190.8, 269.2, 97.7, 103.2]
+patched_conn = [67.9, 148.7, 188.8, 190.9]
+
+fig, ax = plt.subplots()
+
+bar_standard = ax.bar([x/5 + 0.15 for x in Xs], standard_conn, 0.35, color='r')
+bar_patched = ax.bar([x/5 + 0.5 for x in Xs], patched_conn, 0.35, color='b')
+
+ax.set_ylabel('Requests/sec (more is better)')
+ax.set_xlabel('Number of Requests')
+ax.set_title('Latency by number of requests')
+ax.legend((bar_standard, bar_patched), ('Standard GC', 'Patched GC'))
+ax.set_xticks([x/5 + 0.5 for x in Xs])
+ax.set_xticklabels([str(x) + 'k' for x in Xs])
+
+fig.tight_layout()
+fig.savefig('chart_latency.pdf')
+fig.savefig('chart_latency.pgf')
+plt.show()

### 2018-03-23 test_env/latency_test.py

@@ -1,0 +1,11 @@
+import subprocess
+
+def do_test(n):
+    print "make start && sleep 10 && httperf --port=4000 --num-conns={} > 4m_with_{}kreq_8g.httperf".format(n, n // 1000)
+
+    subprocess.check_output(
+        "make start && sleep 10 && httperf --port=4000 --num-conns={} > 4m_with_{}kreq_8g.httperf".format(n, n // 1000),
+        shell=True)
+
+for n in [5000, 10000, 15000, 20000]:
+    do_test(n)

### 2018-03-23 test_env/mysite/mysite/settings.py

@@ -120,9 +120,13 @@ USE_TZ = True
 STATIC_URL = '/static/'
 STATIC_ROOT = os.path.join(BASE_DIR, "static/")
 
-lists = []
-strs = []
-for i in range(16000):
-    lists.append(['poop'] * i)
-    for j in range(40):
-        strs.append('lol' * 8)
+class LargeObject():
+    def __init__(self):
+        self.lists = []
+        self.strs = []
+        for i in range(16000):
+            self.lists.append(['poop'] * i)
+            for j in range(40):
+                self.strs.append('lol' * 8)
+
+static_memory = [LargeObject() for x in range(5)]

### 2018-03-23 test_env/run_test.py

@@ -46,7 +46,7 @@ def measure_all_pids():
         except:
             print "Couldn't get measurement for {} at {}".format(pid, num_requests)
 
-def test_load(requests=30):
+def test_load(requests=50):
     global num_requests
     num_requests += requests * 2
 
@@ -94,7 +94,7 @@ def plot_results(measurements):
 
 if __name__ == '__main__':
     print sys.argv[1]
-    for _ in range(60):
+    for _ in range(50):
         measure_all_pids()
         test_load()
         print _

### 2018-03-23 test_env/show_graph.py

@@ -1,0 +1,55 @@
+import matplotlib.pyplot as plt
+import sys
+import json
+from matplotlib2tikz import save as tikz_save
+
+def pid_to_color(pid):
+    colors = ['r', 'g', 'b', 'k', 'c', 'purple']
+    return colors[int(pid) % len(colors)]
+
+def plot_results(measurements):
+    fig, (shared, private) = plt.subplots(2, 1)
+
+    Xs = []
+    shared_Ys = []
+    private_Ys = []
+    for pid, worker in measurements.items():
+        if pid in ["18622", "14655"]:
+            continue
+        for x, vals in worker.items():
+            if x == "0":
+                continue
+            Xs.append(x)
+            shared_Ys.append((vals['shared'] / 1000.0, pid_to_color(pid)))
+            private_Ys.append((vals['private'] / 1000.0, pid_to_color(pid)))
+
+    shared.scatter(Xs, [x[0] for x in shared_Ys], color=[x[1] for x in shared_Ys])
+    shared.set_xlabel("Requests Served")
+    shared.set_ylabel("Memory (Mb)")
+    shared.set_title("Shared memory")
+    shared.set_xlim(left=0, right=5000)
+    shared.set_ylim(bottom=0, top=6000)
+    shared.get_yaxis().get_major_formatter().set_useOffset(False)
+
+    private.scatter(Xs, [x[0] for x in private_Ys], color=[x[1] for x in private_Ys])
+    private.set_xlabel("Requests Served")
+    private.set_ylabel("Memory (Mb)")
+    private.set_title("Nonshared memory")
+    private.set_xlim(left=0, right=5000)
+    private.set_ylim(bottom=0, top=1400)
+
+    fig.tight_layout()
+    fig.savefig('standard_mem.pdf')
+    fig.savefig('standard_mem.pgf')
+    plt.show()
+
+    """
+    tikz_save('patched_mem.tikz',
+              figureheight = '\\figureheight',
+              figurewidth = '\\figurewidth')
+    """
+
+if __name__ == '__main__':
+	with open(sys.argv[1]) as infile:
+		data = json.load(infile)
+		plot_results(data)
